---
title: "Zero Value Inclusion"
author: "Dana Roemling"
date: '2023-07-25'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Zero Value

So far, when filtering, I have discarded locations with zero value. But since they do tell me something about the location, they should be included from now on, so this RMD reflects the kriging and mapping code including zero. 

```{r library and setup}
# libraries
library(dplyr)
library(spdep)
library(sf)
library(sp)
library(tidyverse)
library(gstat)
library(stringr) 
library(scales)
library(classInt)
library(viridis)
library(viridisLite)
library(rnaturalearth)
library(sfheaders)

# data
longlat <- read.csv(file = './data_maps/gsa_geo_filtered.csv')
longlat_only <- longlat %>% dplyr::select(lon, lat)
corpus <- read.csv(file = './data_ling/full_matrix_for_filtering.csv') 
token_at_location <- read.csv(file = './data_ling/tokens_at_location.csv')
colnames(token_at_location) <- c("City", "Tokencount") 
longlat_small <- read.csv(file = './data_maps/gsa_geo_filtered_nosmall22.csv')
longlat_small <- read.csv(file = './data_maps/gsa_geo_filtered_nosmall85.csv')

# set up mapping & kriging
cities <- data.frame(
  City = c("Köln", "München", "Wien", "Zürich", "Berlin", "Hamburg"),
  Long = c(6.9578, 11.5755, 16.3731, 8.5417, 13.3833, 10),
  Lat = c(50.9422, 48.1372, 48.2083, 47.3769, 52.5167, 53.55))
crs2 <- CRS("+init=epsg:4326")
cities_sf <- st_as_sf(cities, coords = c("Long", "Lat"), crs = crs2)

EU <- st_read(dsn="./data_maps/EU/Try/", layer="NUTS_RG_20M_2021_3035")
EU <- st_transform(EU, CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"))
gsa_outline <- ne_countries(country = c("Austria", "Germany", "Switzerland"), returnclass="sf", scale = "large")
gsa_plot <- gsa_outline %>% dplyr::select(geometry) # just the outlines of the three
gsa_spatial <- as_Spatial(gsa_plot) # new object as spatial instead of sf
sp_grid <- as.data.frame(spsample(gsa_spatial,
                                  n = 100000,
                                  type = "regular", # systematically aligned sampling
                                  offset = c(0.5,0.5))) # makes sure grid is the same every time
sp_grid_sf <- st_as_sf(sp_grid, coords=c("x1","x2"), crs = st_crs(4326))
sp_grid_small <- as.data.frame(spsample(gsa_spatial,
                                        n = 500,
                                        type = "regular", # systematically aligned sampling
                                        offset = c(0.5,0.5))) # makes sure grid is the same every time
sp_grid_sf_small <- st_as_sf(sp_grid_small, coords=c("x1","x2"), crs = st_crs(4326))
```

## Data

So, now when I filter, instead of just adding the long lat data in, I add the filtered data to the long lat introducing 0 for unobserved.

```{r data}
# get word
one_word <- corpus %>% filter(word == "danke")
colnames(one_word) <- c("Token", "City", "Frequency")

# join to all longlat and introduce 0 for NA
new_join <- merge(longlat_small, one_word, by.x ="City", by.y = "City", all = TRUE)
new_join <- new_join %>% dplyr::select("City", "lon", "lat", "Frequency")
new_join[is.na(new_join)] <- 0

# do the relative frequency
merger <- merge(new_join, token_at_location, by.x ="City", by.y = "City")
merger$relfreq <- (merger$Frequency/merger$Tokencount)
merger$relfreq1000 <- (merger$relfreq*1000)
colnames(merger) <- c("City", "lon", "lat", "Frequency", "TokenCount", "RelativeFrequency", "RelativeFrequencyThousand")
summary(merger$TokenCount)
```

## Visualisation

Kriging, Variogram and the whole thing

```{r stats}
trial <- merger %>% dplyr::select("lon", "lat", "RelativeFrequencyThousand") 
data <- trial[3]
coords <- trial[1:2]
crs <- CRS("+init=epsg:4326") 
spdf <- SpatialPointsDataFrame(coords      = coords,
                               data        = data, 
                               proj4string = crs)
spdf = spdf[which(!duplicated(spdf@coords)), ]
vg <- variogram(RelativeFrequencyThousand~1, data = spdf, width = 1, cutoff = 300)
vg_fit <- fit.variogram(vg, vgm("Exp")) 

krig2 <- krige(RelativeFrequencyThousand~1, 
               locations = spdf, 
               newdata = sp_grid_sf_small, 
               model = vg_fit)

krig3 <- krige(var1.pred~1, 
               locations = krig2, 
               newdata = sp_grid_sf, 
               model = vg_fit)
```

## Mapping

```{r map}
legend_max <- round(max(krig3$var1.pred), digits = 2)
legend_min <- round(min(krig3$var1.pred), digits = 2)

ggplot() +
  geom_sf(data = krig3, aes(fill=var1.pred), shape = 21, size = 1, stroke = 0, lwd = 0) +
  geom_sf(data = gsa_plot, aes(geometry = geometry), color="black", fill=NA, size = 0.5) +
  geom_sf_text(data = cities_sf, aes(label = City), size = 2.5, nudge_x = 0, nudge_y = -0.15, family = "Optima") +
  geom_sf(data = cities_sf, aes(geometry = geometry), shape = 4) +
  theme_minimal() +
  scale_fill_steps(low="#ffffff", high="#310d59", n.breaks = 30, labels = NULL) +
  theme(axis.title.x = element_blank(), 
        axis.title.y = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        panel.grid.major = element_blank()) +
  theme(legend.position = c(0.90, 0.65), 
        legend.title = element_text(size = 6, family = "Optima"), 
        legend.text = element_text(size = 6, family = "Optima", hjust = 1),
        legend.title.align = 0,
        legend.key.size = unit(0.3, "cm"),
        legend.key.width = unit(0.4,"cm")) +
  annotate(geom="text", x = 16.6, y = 52.4, label=legend_max, size = 1.5, family = "Optima") +
  annotate(geom="text", x = 16.6, y = 51.3, label=legend_min, size = 1.5, family = "Optima") +
  annotate(geom="text", x = 16.2, y = 49.8, label="danaroemling.com", size = 2.5, family = "Optima") +
  annotate(geom="text", x = 16.2, y = 49.6, label="06/12/2023", size = 2.5, family = "Optima") +
  labs(fill = "Brötchen") 
```