---
title: "Maps from 10k Workflow"
author: "Dana"
date: "2023-12-04"
output:
  html_document: default
  pdf_document: default
toc: true
toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Creating maps with n words from 10k dataframe

This markdown has the steps for the workflow to create maps from already kriged values.

First, I set up the space.
Step 1: Libraries

```{r libs}
# Libraries
library(dplyr)          # For data manipulation and transformation 
library(spdep)          # For spatial dependence, including calculations 
                        # like Moran's I and spatial weights.
library(sf)             # For handling spatial vector data using SF
library(sp)             # For classes and methods for spatial data in R
library(tidyverse)      # A collection of R packages for data science
library(gstat)          # For geostatistical modeling, 
                        # including kriging and variogram analysis.
library(stringr)        # For string manipulation and regular expression operations
library(rnaturalearth)  # For accessing and visualizing natural earth data
```

Step 2: Getting the data

```{r data}
# Get the kriged & morans data
# Load Moran's I data from a CSV file, specifying the path to the file
morans <- read.csv(file = 
                     '/rds/projects/g/grievej-german-dialect-profiling/kriged/morans_10k.csv', 
                   header = TRUE) 

# Convert the data to a data frame
morans <- as.data.frame(morans) 

# Rename the columns for easier reference
colnames(morans) <- c("word", 
                      "score") 

# Assign sequential row names (1 to number of rows)
rownames(morans) <- 1:nrow(morans)

# Load kriged values from a CSV file, specifying the path to the file
kriged <- read.csv(file = '/rds/projects/g/grievej-german-dialect-profiling/kriged/kriged_10k_300cutoff.csv', 
                   header = TRUE) 

# Remove periods from column names for cleaner formatting
names(kriged) <- gsub("\\.", 
                      "", 
                      names(kriged))

# Remove the prefix 'X' from column names, which may indicate that the column 
# names were originally invalid R identifiers
names(kriged) <- gsub("\\<X", 
                      "", 
                      names(kriged))

# Note: The current dataset contains 9250 kriged values until further updates are made
```

Step 3: Set up mapping
This is just all the underlying information needed to produce maps later on.

```{r mapping_basics}
# Set up mapping 
# Create a data frame containing city names and their corresponding 
# longitude and latitude coordinates
cities <- data.frame(
  City = c("Köln", "München", "Wien", "Zürich", "Berlin", "Hamburg"),  # City names
  Long = c(6.9578, 11.5755, 16.3731, 8.5417, 13.3833, 10),  # Longitude values
  Lat = c(50.9422, 48.1372, 48.2083, 47.3769, 52.5167, 53.55))  # Latitude values

# Define the coordinate reference system (CRS) using EPSG:4326 (WGS 84)
crs2 <- CRS("+init=epsg:4326")

# Convert the cities data frame to a simple features (sf) object, 
# specifying the coordinates and CRS
cities_sf <- st_as_sf(cities, 
                      coords = c("Long", 
                                 "Lat"), 
                      crs = crs2)

# Retrieve country outlines for Austria, Germany, and Switzerland 
# as a simple features object
gsa_outline <- ne_countries(country = c("Austria", 
                                        "Germany", 
                                        "Switzerland"), 
                            returnclass="sf", 
                            scale = "large")

# Select only the geometry column from the country outlines for plotting
gsa_plot <- gsa_outline %>% 
  dplyr::select(geometry)

# Convert the sf object to a spatial object for compatibility with sp functions
gsa_spatial <- as_Spatial(gsa_plot)

# Generate a spatial grid by systematically sampling points within the outline
sp_grid <- as.data.frame(spsample(gsa_spatial,
                                  n = 100000,  
                                  # Number of points to sample
                                  type = "regular",  
                                  # Systematically aligned sampling
                                  offset = c(0.5, 0.5)))  
                                  # Offset to ensure a consistent grid layout

# Convert the generated grid to a simple features (sf) object
sp_grid_sf <- st_as_sf(sp_grid, 
                       coords=c("x1", 
                                "x2"), 
                       crs = st_crs(4326))

# Generate a smaller spatial grid by systematically sampling 
# fewer points within the outlined area
sp_grid_small <- as.data.frame(spsample(gsa_spatial,
                                        n = 500,  
                                        # Reduced number of points for the smaller grid
                                        type = "regular",  
                                        # Systematically aligned sampling
                                        offset = c(0.5, 0.5)))  
                                        # Offset for consistent grid layout

# Convert the smaller grid to a simple features (sf) object
sp_grid_sf_small <- st_as_sf(sp_grid_small, 
                             coords=c("x1", 
                                      "x2"), 
                             crs = st_crs(4326))
```

## Unweighted Maps

To trial the code, I first produce maps that are unweighted.
First, I create a variable for the strings I want to work with. Then I get
the kriged values for these.

```{r get_kriged_values}
# Input the words you want to map
# Define a character vector of words, converting them to lowercase
strings <- tolower(c("schlaf", "da", "gehen", "auch", "folgen", "aus"))

# Alternatively, read words from a text file, converting them to lowercase
# Each word should be separated by a comma in the file
strings <- tolower(readLines("/rds/projects/g/grievej-german-dialect-profiling/liste.txt"))

# Trim whitespace and split the input string by commas, 
# unlisting it to create a character vector
strings <- trimws(unlist(strsplit(strings, 
                                  ",")))

# Get kriged values for those words
# Convert the kriged data frame to a tibble for easier manipulation
kriged_tibble <- as_tibble(kriged)

# Select only the longitude and latitude columns from the kriged tibble
kriged_values <- kriged_tibble %>% 
  select(c("lon", 
           "lat"))

# Loop through each word in the strings vector
for (cname in strings) {
  # Check if the current word (cname) is a column in the kriged tibble
  if (cname %in% colnames(kriged_tibble)) {
    # If the word exists, add its corresponding kriged values 
    # to the kriged_values data frame
    kriged_values[, paste0(cname)] <- as.numeric(unlist(kriged_tibble[, cname]))
  }
}
```

In the next step, I calculate the mean, median and SD for the kriged
values in question. 

```{r kriged_calc}
# The 'kriged_values' data frame contains the kriged values for specified words 
# along with their coordinates (lon, lat).

# Select all columns from 'kriged_values' 
# except 'lon' and 'lat' for further calculations
data_for_calculations <- kriged_values %>%
  select(-lon, 
         -lat)

# Calculate the mean of the values for each row in 'data_for_calculations'
# The mean is added as a new column named 'mean' in the 'kriged_values' df
kriged_values$mean <- rowMeans(data_for_calculations)

# Calculate the median for each row in 'data_for_calculations'
# The median is added as a new column named 'median' in the 'kriged_values' df
kriged_values$median <- apply(data_for_calculations, 
                              1, 
                              median)

# Calculate the standard deviation for each row in 'data_for_calculations'
# The standard deviation is added as a new column named 'sd' in the 'kriged_values' df
kriged_values$sd <- apply(data_for_calculations, 
                          1, 
                          sd)

# Convert the resulting 'kriged_values' data frame into a spatial object
# Using the coordinates (lon, lat) with the specified CRS 
calculation_sf <- st_as_sf(kriged_values, 
                           coords = c("lon", 
                                      "lat"), 
                           crs = CRS("+init=epsg:4326"))
```

I can already plot this, if I wanted to, even though this is not as informative.

```{r plot_no_large_grid}
# Start a new ggplot object
ggplot() +
    # Add a spatial polygon layer representing kriged mean values
    geom_sf(data = calculation_sf, 
            aes(fill = mean_val), 
            shape = 21, 
            size = 2, 
            stroke = 0, 
            lwd = 0) +
    
    # Overlay the outline of the geographical area (gsa_plot) with black borders
    geom_sf(data = gsa_plot, 
            aes(geometry = geometry), 
            color = "black", 
            fill = NA, 
            size = 0.5) +
    
    # Add labels for cities from the cities_sf data, 
    # positioned slightly below each city
    geom_sf_text(data = cities_sf, 
                 aes(label = City), 
                 size = 2.5, 
                 nudge_x = 0, 
                 nudge_y = -0.15, 
                 family = "Optima") +
    
    # Add points for cities using a cross (shape 4)
    geom_sf(data = cities_sf, 
            aes(geometry = geometry), 
            shape = 4) +
    
    # Use a minimal theme for the plot
    theme_minimal() +
    
    # Define the fill color scale for the mean values, 
    # transitioning from white to a dark color
    scale_fill_steps(low = "#ffffff", 
                     high = "#310d59", 
                     n.breaks = 20, 
                     labels = NULL) +
    
    # Customize theme to remove axis titles and text, and remove major grid lines
    theme(axis.title.x = element_blank(), 
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          panel.grid.major = element_blank()) +
    
    # Customize legend appearance
    # Position the legend
    theme(legend.position = c(0.90, 
                              0.65),  
          # Font size and family for legend title
          legend.title = element_text(size = 6, 
                                      family = "Optima"),  
          # Font size and alignment for legend text
          legend.text = element_text(size = 6, 
                                     family = "Optima",
                                     hjust = 1), 
          # Align the legend title to the left
          legend.title.align = 0,  
          # Size of the legend keys
          legend.key.size = unit(0.3, 
                                 "cm"),  
          # Width of the legend keys
          legend.key.width = unit(0.4, 
                                  "cm")) +  
    
    # Optional annotations for the legend 
    # ggplot2::annotate(geom = "text", 
    #                   x = 17, 
    #                   y = 52.4, 
    #                   label = legend_max, 
    #                   size = 1.5, 
    #                   family = "Optima") +
  
    # ggplot2::annotate(geom = "text", 
    #                   x = 17, 
    #                   y = 51.3, 
    #                   label = legend_min, 
    #                   size = 1.5, 
    #                   family = "Optima") +
    
    # Add a label for the fill legend (currently left blank)
    labs(fill = " ")
```

So, I rather krige to the large grid before mapping.

```{r large_krige_one}
# Variogram
# Extract the mean values from the kriged_values dataframe for analysis
data <- kriged_values["mean"]

# Get the coordinates (longitude and latitude) for the spatial points
coords <- kriged_values[c("lon", 
                          "lat")]

# Define the coordinate reference system (CRS) for geographic coordinates (WGS 84)
crs <- CRS("+init=epsg:4326") 

# Create a SpatialPointsDataFrame object to hold the spatial data 
spdf <- SpatialPointsDataFrame(coords      = coords,  # Coordinates for spatial points
                               data        = data,   # Associated data (mean values)
                               proj4string = crs)    # Assign the defined CRS

# Remove duplicated spatial coordinates to avoid redundancy in the analysis
spdf = spdf[which(!duplicated(spdf@coords)), ]

# Compute the empirical variogram for the mean values; 
# 'mean~1' indicates a constant mean model, 'width' defines bin width, 
# 'cutoff' is the distance limit
vg <- variogram(mean ~ 1, 
                data = spdf, 
                width = 1, 
                cutoff = 300)

# Fit a variogram model to the empirical variogram 
# using an exponential model ('Exp')
vg_fit <- fit.variogram(vg, 
                        vgm("Exp")) 

# Kriging
# Perform ordinary kriging interpolation for the mean values
# 'mean~1' indicates a constant mean model, 'locations' are the spatial points, 
# 'newdata' specifies the grid for prediction,
# and 'model' is the fitted variogram model
krig <- krige(mean ~ 1, 
               locations = spdf, 
               newdata = sp_grid_sf, 
               model = vg_fit)
```

This then can be turned into a plot.

```{r plot_one}
# Calculate the maximum and minimum predicted values from the kriging results
# Round the maximum predicted value to two decimal places
legend_max <- round(max(krig$var1.pred), 
                    digits = 2)  
legend_min <- round(min(krig$var1.pred), 
                    digits = 2)  

# Create a ggplot visualization
ggplot() +
    # Plot the kriging predictions as a filled spatial polygon layer
    geom_sf(data = krig, 
            aes(fill = var1.pred), 
            shape = 21, 
            size = 0.75, 
            stroke = 0, 
            lwd = 0) +
    
    # Overlay the outlines of the geographical areas (GSA outline)
    geom_sf(data = gsa_plot, 
            aes(geometry = geometry), 
            color = "black", 
            fill = NA, 
            size = 0.5) +
    
    # Add city labels from the cities_sf spatial data
    geom_sf_text(data = cities_sf, 
                 aes(label = City), 
                 size = 2.5, 
                 nudge_x = 0, 
                 nudge_y = -0.15, 
                 family = "Optima") +
    
    # Add city points for better visibility
    geom_sf(data = cities_sf, 
            aes(geometry = geometry), 
            shape = 4) +
    
    # Apply a minimal theme for clean visualization
    theme_minimal() +
    
    # Define color scale for the fill aesthetic, from white to dark blue
    scale_fill_steps(low = "#ffffff", 
                     high = "#310d59", 
                     n.breaks = 20, 
                     labels = NULL) +
    
    # Customize the theme to remove axis titles and text for a cleaner look
    theme(axis.title.x = element_blank(), 
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          panel.grid.major = element_blank()) +
    
    # Customize legend appearance and position
    theme(legend.position = c(0.90, 
                              0.65), 
          legend.title = element_text(size = 6, 
                                      family = "Optima"), 
          legend.text = element_text(size = 6, 
                                     family = "Optima", 
                                     hjust = 1),
          legend.title.align = 0,
          legend.key.size = unit(0.3, 
                                 "cm"),
          legend.key.width = unit(0.4, 
                                  "cm")) +
    
    # Annotate the plot with the maximum predicted value
    ggplot2::annotate(geom = "text", 
                      x = 17, 
                      y = 52.4, 
                      label = legend_max, 
                      size = 1.5, 
                      family = "Optima") +
    
    # Annotate the plot with the minimum predicted value
    ggplot2::annotate(geom = "text", 
                      x = 17, 
                      y = 51.3, 
                      label = legend_min, 
                      size = 1.5, 
                      family = "Optima") +
    
    # Set the label for the fill legend to be empty
    labs(fill = " ")
```

## Moran's I weighted maps

Similar calculation as above, but first I multiply all values by their
respective Moran's I value.

```{r morans_calc}
# Multiply kriged values by Moran's I value for each specified word
# Add the resulting values as new columns in the kriged_values dataframe
for (word in strings) {
  # Check if the current word is present in the morans dataframe
  if (word %in% morans$word) {
    # Extract the corresponding Moran's I score for the word
    word_score <- as.numeric(morans$score[morans$word == word])
    
    # Print the word, its score, and its index for debugging purposes
    print(paste0("Word: ", word))
    print(paste0("Score: ", word_score))
    print(paste0("Index: ", match(word, 
                                  morans$word)))
    
    # Multiply the kriged values for the word by its corresponding Moran's I score
    kriged_values[, paste0(word, "_multiplied")] <- 
      as.numeric(unlist(kriged_values[, word])) * word_score
  }
}

# Calculate summary statistics (mean, median, and standard deviation) 
# for the multiplied columns
# Create a vector of the new column names
multiplied_columns <- paste0(strings, 
                             "_multiplied")  

calculation <- kriged_values %>%
  # Select the longitude, latitude, and multiplied columns
  dplyr::select(lon, 
                lat, 
                all_of(multiplied_columns))  

# Compute the mean across the multiplied columns, ignoring NA values
calculation$mean <- rowMeans(calculation[, multiplied_columns], 
                             na.rm = TRUE)

# Compute the median across the multiplied columns, ignoring NA values
calculation$median <- apply(calculation[, multiplied_columns], 
                            1, median, 
                            na.rm = TRUE)

# Compute the standard deviation across the multiplied columns, 
# ignoring NA values
calculation$sd <- apply(calculation[, multiplied_columns], 
                        1, 
                        sd, 
                        na.rm = TRUE)

# Convert the results to a simple features (sf) object 
calculation_sf <- st_as_sf(calculation, 
                           coords = c("lon",
                                      "lat"), 
                           crs = crs2)
```

Then, again, kriging to large grid.

```{r large_krige_two}
# Variogram Analysis
# Extract the 'mean' column from the calculation dataframe
data <- calculation["mean"]  
# Extract longitude and latitude coordinates
coords <- calculation[c("lon", 
                        "lat")]  
# Define the coordinate reference system (EPSG:4326 for WGS 84)
crs <- CRS("+init=epsg:4326") 

# Create a SpatialPointsDataFrame object combining coordinates and data and
# assign the specified CRS to the spatial object
spdf <- SpatialPointsDataFrame(coords      = coords,
                               data        = data, 
                               proj4string = crs) 

# Remove duplicate coordinates from the SpatialPointsDataFrame
spdf = spdf[which(!duplicated(spdf@coords)), ]

# Compute the variogram for the mean values with specified parameters
vg <- variogram(mean~1, 
                data = spdf, 
                width = 1, 
                cutoff = 300) 
# 'mean~1' specifies that we're modeling the mean values without a trend
# 'width' sets the distance interval for calculating the variogram
# 'cutoff' sets the maximum distance for which the variogram is calculated

# Fit an exponential variogram model to the computed variogram
vg_fit <- fit.variogram(vg, 
                        vgm("Exp")) 
# 'vgm("Exp")' specifies that we're using an exponential model for fitting

# Kriging interpolation based on the fitted variogram model
krig <- krige(mean~1,  # Model the mean values
               locations = spdf,  # Use the spatial points from the spdf
               newdata = sp_grid_sf,  # Predict values at the points in sp_grid_sf
               model = vg_fit)  # Use the fitted variogram model for the kriging
```

Then, this can be mapped as well. 

```{r plot_two}
# Calculate maximum and minimum predicted values from the kriging results
# Round the maximum predicted value to 2 decimal places
legend_max <- round(max(krig$var1.pred), 
                    digits = 2)  
legend_min <- round(min(krig$var1.pred), 
                    digits = 2) 

# Create a ggplot visualization
ggplot() +
    # Add the kriging results as a filled map
    geom_sf(data = krig, 
            aes(fill = var1.pred), 
            shape = 21, 
            size = 0.75, 
            stroke = 0, 
            lwd = 0) +
  
    # Add the geographical outline of the selected countries
    geom_sf(data = gsa_plot, 
            aes(geometry = geometry), 
            color="black", 
            fill = NA, 
            size = 0.5) +
  
    # Add city names to the map, nudging the text slightly for better visibility
    geom_sf_text(data = cities_sf, 
                 aes(label = City), 
                 size = 2.5, 
                 nudge_x = 0, 
                 nudge_y = -0.15, 
                 family = "Optima") +
  
    # Add city markers to the map
    geom_sf(data = cities_sf, 
            aes(geometry = geometry), 
            shape = 4) +
  
    # Set a minimal theme for the plot
    theme_minimal() +
  
    # Define a gradient color scale for the fill based on predicted values
    scale_fill_steps(low = "#ffffff", 
                     high = "#310d59", 
                     n.breaks = 20, 
                     labels = NULL) +
  
    # Customize the appearance of axis titles and text
    theme(axis.title.x = element_blank(), 
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          panel.grid.major = element_blank()) +
  
    # Customize the legend appearance
    # Position of the legend in the plot
    theme(legend.position = c(0.90, 
                              0.65),  
          # Title text settings
          legend.title = element_text(size = 6, 
                                      family = "Optima"),  
          # Legend item text settings
          legend.text = element_text(size = 6, 
                                     family = "Optima", 
                                     hjust = 1),  
          legend.title.align = 0,  # Align legend title to the left
          legend.key.size = unit(0.3, 
                                 "cm"),  # Size of the legend keys
          legend.key.width = unit(0.4,
                                  "cm")) +  # Width of the legend keys
  
    # Annotate the plot with maximum and minimum values for reference
    ggplot2::annotate(geom = "text", 
                      x = 17, 
                      y = 52.4, 
                      label = legend_max, 
                      size = 1.5, 
                      family = "Optima") +
  
    ggplot2::annotate(geom = "text", 
                      x = 17, 
                      y = 51.3, 
                      label = legend_min, 
                      size = 1.5, 
                      family = "Optima") +
  
    # Remove the default fill label for the legend
    labs(fill = " ")
```
