---
title: "Maps from 10k Workflow"
author: "Dana"
date: "2023-12-04"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Creating maps with n words from 10k dataframe

This markdown has the steps for the workflow to create maps from already kriged values.

First, I set up the space.
Step 1: Libraries

```{r libs}
# Libraries
library(dplyr)
library(spdep)
library(sf)
library(sp)
library(tidyverse)
library(gstat)
library(stringr) 
library(rnaturalearth)
```

Step 2: Getting the data

```{r data}
# Get the kriged & morans data
morans <- read.csv(file = 
                     '/rds/projects/g/grievej-german-dialect-profiling/kriged/morans_10k.csv', 
                   header = TRUE) 
morans <- as.data.frame(morans)
colnames(morans) <- c("word", "score") 
rownames(morans) <- 1:nrow(morans)

kriged <- read.csv(file = 
                     '/rds/projects/g/grievej-german-dialect-profiling/kriged/kriged_10k.csv', 
                   header = TRUE) 
names(kriged) <- gsub("\\.", "", names(kriged))
names(kriged) <- gsub("\\<X", "", names(kriged))
# for now this is 9250 kriged values until this is rerun
```

Step 3: Set up mapping
This is just all the underlying information needed to produce maps later on.

```{r mapping_basics}
# set up mapping 
cities <- data.frame(
  City = c("Köln", "München", "Wien", "Zürich", "Berlin", "Hamburg"),
  Long = c(6.9578, 11.5755, 16.3731, 8.5417, 13.3833, 10),
  Lat = c(50.9422, 48.1372, 48.2083, 47.3769, 52.5167, 53.55))
crs2 <- CRS("+init=epsg:4326")
cities_sf <- st_as_sf(cities, coords = c("Long", "Lat"), crs = crs2)

gsa_outline <- ne_countries(country = c("Austria", "Germany", "Switzerland"), returnclass="sf", scale = "large")
gsa_plot <- gsa_outline %>% dplyr::select(geometry) # just the outlines of the three
gsa_spatial <- as_Spatial(gsa_plot) # new object as spatial instead of sf

sp_grid <- as.data.frame(spsample(gsa_spatial,
                                  n = 100000,
                                  type = "regular", # systematically aligned sampling
                                  offset = c(0.5,0.5))) # makes sure grid is the same every time
sp_grid_sf <- st_as_sf(sp_grid, coords=c("x1","x2"), crs = st_crs(4326))

sp_grid_small <- as.data.frame(spsample(gsa_spatial,
                                        n = 500,
                                        type = "regular", # systematically aligned sampling
                                        offset = c(0.5,0.5))) # makes sure grid is the same every time
sp_grid_sf_small <- st_as_sf(sp_grid_small, coords=c("x1","x2"), crs = st_crs(4326))
```

## Unweighted Maps

To trial the code, I first produce maps that are unweighted.
First, I create a variable for the strings I want to work with. Then I get
the kriged values for these.

```{r get_kriged_values}
# input the words you want to map
strings <- tolower(c("schlaf", "da","gehen","auch","folgen","aus"))
# OR
strings <- tolower(readLines("/rds/projects/g/grievej-german-dialect-profiling/liste.txt"))
strings <- trimws(unlist(strsplit(strings, ",")))

# get kriged values for those words
kriged_tibble <- as_tibble(kriged)
kriged_values <- kriged_tibble %>% select(c("lon", "lat"))

for (cname in strings) {
  if (cname %in% colnames(kriged_tibble)) {
    kriged_values[, paste0(cname)] <- as.numeric(unlist(kriged_tibble[, cname]))
  }
}

```

In the next step, I calculate the mean, median and SD for the kriged
values in question. 

```{r kriged_calc}
# these values were scaled during the kriging process using scale() 
calculation <- kriged_values %>%
  dplyr::rowwise() %>%
  dplyr::mutate(mean_val = mean(c_across(strings), na.rm = TRUE),
                median_val = median(c_across(strings), na.rm = TRUE),
                sd_val = sd(c_across(strings), na.rm = TRUE)) %>%
  dplyr::ungroup()
calculation_sf <- st_as_sf(calculation, coords = c("lon", "lat"), crs = CRS("+init=epsg:4326"))
```

I can already plot this, if I wanted to, even though this is not as informative.

```{r plot_no_large_grid}
ggplot() +
    geom_sf(data = calculation_sf, aes(fill=mean_val), shape = 21, size = 2, stroke = 0, lwd = 0) +
    geom_sf(data = gsa_plot, aes(geometry = geometry), color="black", fill=NA, size = 0.5) +
    geom_sf_text(data = cities_sf, aes(label = City), size = 2.5, nudge_x = 0, nudge_y = -0.15, family = "Optima") +
    geom_sf(data = cities_sf, aes(geometry = geometry), shape = 4) +
    theme_minimal() +
    scale_fill_steps(low="#ffffff", high="#310d59", n.breaks = 20, labels = NULL) +
    theme(axis.title.x = element_blank(), 
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          panel.grid.major = element_blank()) +
    theme(legend.position = c(0.90, 0.65), 
          legend.title = element_text(size = 6, family = "Optima"), 
          legend.text = element_text(size = 6, family = "Optima", hjust = 1),
          legend.title.align = 0,
          legend.key.size = unit(0.3, "cm"),
          legend.key.width = unit(0.4,"cm")) +
    #ggplot2::annotate(geom="text", x = 17, y = 52.4, label=legend_max, size = 1.5, family = "Optima") +
    #ggplot2::annotate(geom="text", x = 17, y = 51.3, label=legend_min, size = 1.5, family = "Optima") +
    labs(fill = " ")
```

So, I rather krige to the large grid before mapping.

```{r large_krige_one}
# Variogram
data <- calculation["mean_val"]
coords <- calculation[c("lon", "lat")]
crs <- CRS("+init=epsg:4326") 
spdf <- SpatialPointsDataFrame(coords      = coords,
                               data        = data, 
                               proj4string = crs)
spdf = spdf[which(!duplicated(spdf@coords)), ]
vg <- variogram(mean_val~1, data = spdf, width = 1, cutoff = 300)
vg_fit <- fit.variogram(vg, vgm("Exp")) 

# Kriging
krig <- krige(mean_val~1, 
               locations = spdf, 
               newdata = sp_grid_sf, 
               model = vg_fit)
```

This then can be turned into a plot.

```{r plot_one}
legend_max <- round(max(krig$var1.pred), digits = 2)
legend_min <- round(min(krig$var1.pred), digits = 2) 

ggplot() +
    geom_sf(data = krig, aes(fill=var1.pred), shape = 21, size = 0.75, stroke = 0, lwd = 0) +
    geom_sf(data = gsa_plot, aes(geometry = geometry), color="black", fill=NA, size = 0.5) +
    geom_sf_text(data = cities_sf, aes(label = City), size = 2.5, nudge_x = 0, nudge_y = -0.15, family = "Optima") +
    geom_sf(data = cities_sf, aes(geometry = geometry), shape = 4) +
    theme_minimal() +
    scale_fill_steps(low="#ffffff", high="#310d59", n.breaks = 20, labels = NULL) +
    theme(axis.title.x = element_blank(), 
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          panel.grid.major = element_blank()) +
    theme(legend.position = c(0.90, 0.65), 
          legend.title = element_text(size = 6, family = "Optima"), 
          legend.text = element_text(size = 6, family = "Optima", hjust = 1),
          legend.title.align = 0,
          legend.key.size = unit(0.3, "cm"),
          legend.key.width = unit(0.4,"cm")) +
    ggplot2::annotate(geom="text", x = 17, y = 52.4, label=legend_max, size = 1.5, family = "Optima") +
    ggplot2::annotate(geom="text", x = 17, y = 51.3, label=legend_min, size = 1.5, family = "Optima") +
    labs(fill = " ")
```

## Moran's I weighted maps

Similar calculation as above, but first I multiply all values by their
respective Moran's I value.

```{r morans_calc}
# calculate kriged values by moran's I value for variable
# add as new columns
for (word in strings) {
  if (word %in% morans$word) {
    word_score <- as.numeric(morans$score[morans$word == word])
    print(paste0("Word: ", word))
    print(paste0("Score: ", word_score))
    print(paste0("Index: ", match(word, morans$word)))
    kriged_values[, paste0(word, "_multiplied")] <- as.numeric(unlist(kriged_values[, word])) * word_score
  }
}

# calculate mean, median, sd
multiplied_columns <- paste0(strings, "_multiplied")
calculation <- kriged_values %>%
  dplyr::select(lon, lat, all_of(multiplied_columns)) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(mean_val = mean(c_across(multiplied_columns), na.rm = TRUE),
                median_val = median(c_across(multiplied_columns), na.rm = TRUE),
                sd_val = sd(c_across(multiplied_columns), na.rm = TRUE)) %>%
  dplyr::ungroup()
calculation_sf <- st_as_sf(calculation, coords = c("lon", "lat"), crs = crs2)
```

Then, again, kriging to large grid.

```{r large_krige_two}
# Variogram
data <- calculation["mean_val"]
coords <- calculation[c("lon", "lat")]
crs <- CRS("+init=epsg:4326") 
spdf <- SpatialPointsDataFrame(coords      = coords,
                               data        = data, 
                               proj4string = crs)
spdf = spdf[which(!duplicated(spdf@coords)), ]
vg <- variogram(mean_val~1, data = spdf, width = 1, cutoff = 300)
vg_fit <- fit.variogram(vg, vgm("Exp")) 

# Kriging
krig <- krige(mean_val~1, 
              locations = spdf, 
              newdata = sp_grid_sf, 
              model = vg_fit)
```

Then, this can be mapped as well. 

```{r plot_two}
legend_max <- round(max(krig$var1.pred), digits = 2)
legend_min <- round(min(krig$var1.pred), digits = 2)  

ggplot() +
    geom_sf(data = krig, aes(fill=var1.pred), shape = 21, size = 0.75, stroke = 0, lwd = 0) +
    geom_sf(data = gsa_plot, aes(geometry = geometry), color="black", fill=NA, size = 0.5) +
    geom_sf_text(data = cities_sf, aes(label = City), size = 2.5, nudge_x = 0, nudge_y = -0.15, family = "Optima") +
    geom_sf(data = cities_sf, aes(geometry = geometry), shape = 4) +
    theme_minimal() +
    scale_fill_steps(low="#ffffff", high="#310d59", n.breaks = 20, labels = NULL) +
    theme(axis.title.x = element_blank(), 
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          panel.grid.major = element_blank()) +
    theme(legend.position = c(0.90, 0.65), 
          legend.title = element_text(size = 6, family = "Optima"), 
          legend.text = element_text(size = 6, family = "Optima", hjust = 1),
          legend.title.align = 0,
          legend.key.size = unit(0.3, "cm"),
          legend.key.width = unit(0.4,"cm")) +
    ggplot2::annotate(geom="text", x = 17, y = 52.4, label=legend_max, size = 1.5, family = "Optima") +
    ggplot2::annotate(geom="text", x = 17, y = 51.3, label=legend_min, size = 1.5, family = "Optima") +
    labs(fill = " ")
```
