---
title: "Kriging by Country Workflow"
author: "Dana Roemling"
date: '2024-02-09'
output: html_document
toc: true
toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Kriging by Country

The idea for this markdown is to first check which is the most likely country of origin for a text and then use only the data from that country for further analysis. 

```{r setting_up, echo=FALSE, message=FALSE, results='hide', warning=FALSE}
# Libraries
library(dplyr)
library(spdep)
library(sf)
library(sp)
library(tidyverse)
library(gstat)
library(stringr) 
library(rnaturalearth)

# Data
# /rds/projects/g/grievej-german-dialect-profiling/kriged/
kriged_all <- read.csv(file = 
                     'data_ling/kriged/kriged_10k_300cutoff.csv', 
                   header = TRUE) 
names(kriged_all) <- gsub("\\.", "", names(kriged_all))
names(kriged_all) <- gsub("\\<X", "", names(kriged_all))
kriged_all <- kriged_all %>% dplyr::select(1:2002)

kriged_ger <- read.csv(file = 
                     'data_ling/kriged/kriged_5k_300cutoff_germany.csv', 
                   header = TRUE) 
names(kriged_ger) <- gsub("\\.", "", names(kriged_ger))
names(kriged_ger) <- gsub("\\<X", "", names(kriged_ger))
kriged_ger <- kriged_ger %>% dplyr::select(1:2002)

kriged_aus <- read.csv(file = 
                     'data_ling/kriged/kriged_5k_300cutoff_austria.csv', 
                   header = TRUE) 
names(kriged_aus) <- gsub("\\.", "", names(kriged_aus))
names(kriged_aus) <- gsub("\\<X", "", names(kriged_aus))
kriged_aus <- kriged_aus %>% dplyr::select(1:2002)

kriged_swi <- read.csv(file = 
                     'data_ling/kriged/kriged_5k_300cutoff_switzerland.csv', 
                   header = TRUE) 
names(kriged_swi) <- gsub("\\.", "", names(kriged_swi))
names(kriged_swi) <- gsub("\\<X", "", names(kriged_swi))
kriged_swi <- kriged_swi %>% dplyr::select(1:2002)

# Geo stuff
ger_outline <- ne_countries(country = "Germany", returnclass="sf", scale = "large")
ger_plot <- ger_outline %>% dplyr::select(geometry) 
ger_spatial <- as_Spatial(ger_plot) 
aus_outline <- ne_countries(country = "Austria", returnclass="sf", scale = "large")
aus_plot <- aus_outline %>% dplyr::select(geometry) 
aus_spatial <- as_Spatial(aus_plot) 
swi_outline <- ne_countries(country = "Switzerland", returnclass="sf", scale = "large")
swi_plot <- swi_outline %>% dplyr::select(geometry) 
swi_spatial <- as_Spatial(swi_plot) 
ger_states <- ne_states(country = "Germany", returnclass = "sf")
ger_states_plot <- ger_states %>% dplyr::select(geometry)
ger_states_spatial <- as_Spatial(ger_states_plot) 
aus_states <- ne_states(country = "Austria", returnclass = "sf")
aus_states_plot <- aus_states %>% dplyr::select(geometry)
aus_states_spatial <- as_Spatial(aus_states_plot) 
swi_states <- ne_states(country = "Switzerland", returnclass = "sf")
swi_states_plot <- swi_states %>% dplyr::select(geometry)
swi_states_spatial <- as_Spatial(swi_states_plot) 
crs3 <- CRS("+init=epsg:4326") 
crs2 <- st_crs(4326)

cities_ger <- data.frame(
  City = c("Köln", "München", "Berlin", "Hamburg", "Frankfurt", "Kassel"),
  Long = c(6.9578, 11.5755, 13.3833, 10, 8.6821, 9.4797),
  Lat = c(50.9422, 48.1372, 52.5167, 53.55, 50.1109, 51.3127))
cities_ger_sf <- st_as_sf(cities_ger, coords = c("Long", "Lat"), crs = crs2)
sp_grid_ger <- as.data.frame(spsample(ger_spatial,
                                  n = 100000,
                                  type = "regular", # systematically aligned sampling
                                  offset = c(0.5,0.5))) # makes sure grid is the same every time
sp_grid_ger_sf <- st_as_sf(sp_grid_ger, coords=c("x1","x2"), crs = crs2)

cities_aus <- data.frame(
  City = c("Wien", "Graz", "Innsbruck"),
  Long = c(16.3731, 15.4333, 11.3933),
  Lat = c(48.2083, 47.0667, 47.2683))
cities_aus_sf <- st_as_sf(cities_aus, coords = c("Long", "Lat"), crs = crs2)
sp_grid_aus <- as.data.frame(spsample(aus_spatial,
                                  n = 50000,
                                  type = "regular", # systematically aligned sampling
                                  offset = c(0.5,0.5))) # makes sure grid is the same every time
sp_grid_aus_sf <- st_as_sf(sp_grid_aus, coords=c("x1","x2"), crs = crs2)

cities_swi <- data.frame(
  City = c("Zürich", "Bern"),
  Long = c(8.5417, 7.4474),
  Lat = c(47.3769, 46.9480))
cities_swi_sf <- st_as_sf(cities_swi, coords = c("Long", "Lat"), crs = crs2)
sp_grid_swi <- as.data.frame(spsample(swi_spatial,
                                  n = 50000,
                                  type = "regular", # systematically aligned sampling
                                  offset = c(0.5,0.5))) # makes sure grid is the same every time
sp_grid_swi_sf <- st_as_sf(sp_grid_swi, coords=c("x1","x2"), crs = crs2)
```

Set up the data you want to analyse.

```{r strings}
# input the words you want to map
strings <- tolower(c("aber", "eigentlich", "isses", "mir", "echt", "egal", "bin", "auch", "offen", "für", "kreative", "sachen"))
strings <- unique(strings)

strings <- tolower(c("Danke", "Endlich", "mal", "jemand", "der", "aus", "Erfahrung", "spricht", "Wie", "gesagt", "ich", "gehe", "mit", "meinem", "Hund", "zum", "Arzt", "aber", "geht", "ihr", "zum", "Arzt", "wenn", "ihr", "mal", "2", "Tage", "hustet", "Ich", "denke", "nicht"))
strings <- unique(strings)

# OR
strings <- tolower(readLines("/rds/projects/g/grievej-german-dialect-profiling/liste.txt"))
strings <- trimws(unlist(strsplit(strings, ",")))
```



## Determining Max Value Country

First, I need to determine which country, based on the overall/averaged kriged values, contains the max point. 

```{r country_pred_function}
country_prediction <- function(strings) {
  kriged_values <- kriged_all %>% select(c("lon", "lat"))
  
  for (cname in strings) {
    if (cname %in% colnames(kriged_all)) {
    kriged_values[, paste0(cname)] <- as.numeric(unlist(kriged_all[, cname]))
    }
  }
  
  data_for_calculations <- kriged_values %>% select(-lon, -lat)
  calculation <- kriged_values %>% select(lon, lat)
  calculation$mean_val <- rowMeans(data_for_calculations)
  calculation$median_val <- apply(data_for_calculations, 1, median)
  calculation$sd_val <- apply(data_for_calculations, 1, sd)
  
  # get max
  mean_pred <- calculation %>% 
    filter(mean_val > 0) %>% 
    arrange(desc(mean_val)) %>%
    slice (1:1) %>% 
    dplyr::select(mean_val, lon, lat)
  
  median_pred <- calculation %>% 
    filter(median_val > 0) %>% 
    arrange(desc(median_val)) %>%
    slice (1:1) %>% 
    dplyr::select(median_val, lon, lat)
  
  # transform max into sf
  mean_pred_sf <- st_as_sf(mean_pred, coords = c("lon", "lat"), crs = crs2)
  median_pred_sf <- st_as_sf(median_pred, coords = c("lon", "lat"), crs = crs2)  
  
  # check if those are within countries:
  Germany_mean <- ger_outline %>% st_contains(mean_pred_sf) %>% as.matrix()
  Austria_mean <- aus_outline %>% st_contains(mean_pred_sf) %>% as.matrix()
  Switzerland_mean <- swi_outline %>% st_contains(mean_pred_sf) %>% as.matrix()
  Germany_median <- ger_outline %>% st_contains(median_pred_sf) %>% as.matrix()
  Austria_median <- aus_outline %>% st_contains(median_pred_sf) %>% as.matrix()
  Switzerland_median <- swi_outline %>% st_contains(median_pred_sf) %>% as.matrix()
  
  answer <- data.frame(matrix(ncol = 3, nrow = 2))
  colnames(answer) <- c("Germany", "Austria", "Switzerland")
  rownames(answer) <- c("Mean", "Median")
  
  if (Germany_mean == 1) {
    answer[1,1] <- TRUE
    }
  if (Austria_mean == 1) {
    answer[1,2] <- TRUE
    }
  if (Switzerland_mean == 1) {
    answer[1,3] <- TRUE
    }
  if (Germany_median == 1) {
    answer[2,1] <- TRUE
    }
  if (Austria_median == 1) {
    answer[2,2] <- TRUE
    }
  if (Switzerland_mean == 1) {
    answer[2,3] <- TRUE
    }
  return(list(
    answer = answer,
    max_val = mean_pred
  ))
}

# calling function
resulting_prediction <- country_prediction(strings)

# see both results
resulting_prediction$answer
resulting_prediction$max_val
```

### Max Value of Individual Words

To export where each word has its hotspot, this function loops over the string list and gets those max values and their locations. This is before noise is removed, so it might be smart to do this for a country only. As the input here is a df with long lat as the first to columns, application is flexible. 

```{r max_val_per_word}
get_max_values <- function(df) {
  results <- list()
  
  # Loop over the columns starting from the third column 
  for (col_name in names(df)[3:ncol(df)]) {
    
    # Set negative to NA to exclude them
    positive_values <- df[[col_name]]
    positive_values[positive_values <= 0] <- NA
    
    # Find the row index of the maximum value in the current column
    max_index <- which.max(positive_values)
    
    # Check if there are any positive values
    if (!is.na(max_index) && !is.na(positive_values[max_index])) {
    
      # Extract the maximum value
      max_value <- df[max_index, col_name]
      
      # Extract the corresponding longitude and latitude
      lon <- df[max_index, 1]
      lat <- df[max_index, 2]
      
      # Create a result list for the current column
      result <- list(
        word = col_name,
        max_value = max_value,
        longitude = lon,
        latitude = lat
      )
    
      # Append the result to the results list
      results[[col_name]] <- result
    } else {
      
      # If no positive value is found, store NA
      result <- list(
        word = col_name,
        max_value = NA,
        longitude = NA,
        latitude = NA
      ) 
      
      # Append the result to the results list
      results[[col_name]] <- result
    }
  }
  
  # Convert the list of results to a data frame
  results_df <- do.call(rbind, lapply(results, as.data.frame))
  
  return(results_df)
}

# create input for this function as above
kriged_values <- kriged_all %>% select(c("lon", "lat"))
for (cname in strings) {
  if (cname %in% colnames(kriged_all)) {
  kriged_values[, paste0(cname)] <- as.numeric(unlist(kriged_all[, cname]))
  }
}

# input is a dataframe, in which long & lat need to be the first two columns
output_df <- get_max_values(kriged_values)
```


# Country level code

To make sure the maps that are being created do not get influenced by noise, the code below focusses on each country to krige to a large grid.
It also contains code to predict the state per country where the max point lies (and extract the max point).

## Germany

```{r ger_kriging}
# get data for strings
kriged_values_ger <- kriged_ger %>% select(c("lon", "lat"))
for (cname in strings) {
  if (cname %in% colnames(kriged_ger)) {
    kriged_values_ger[, paste0(cname)] <- as.numeric(unlist(kriged_ger[, cname]))
  }
}

# create new list of column names for further operations
strings <- toString(colnames(kriged_values_ger[3:ncol(kriged_values_ger)]))
strings <- trimws(unlist(strsplit(strings, ",")))

data_for_calculations_ger <- kriged_values_ger %>% select(-lon, -lat)
calculation_ger <- kriged_values_ger %>% select(lon, lat)
calculation_ger$mean_val <- rowMeans(data_for_calculations_ger)
calculation_ger$median_val <- apply(data_for_calculations_ger, 1, median)
calculation_ger$sd_val <- apply(data_for_calculations_ger, 1, sd)

calculation_ger_sf <- st_as_sf(calculation_ger, coords = c("lon", "lat"), crs = crs2)

# get max
mean_pred_ger <- calculation_ger %>% 
  filter(mean_val > 0) %>% 
  arrange(desc(mean_val)) %>%
  slice (1:1) %>% 
  dplyr::select(mean_val, lon, lat)

median_pred_ger <- calculation_ger %>% 
  filter(median_val > 0) %>% 
  arrange(desc(median_val)) %>%
  slice (1:1) %>% 
  dplyr::select(median_val, lon, lat)

# Variogram
data_ger <- calculation_ger["mean_val"]
coords_ger <- calculation_ger[c("lon", "lat")]
spdf_ger <- SpatialPointsDataFrame(coords      = coords_ger,
                               data        = data_ger, 
                               proj4string = crs3)
spdf_ger = spdf_ger[which(!duplicated(spdf_ger@coords)), ]
vg_ger <- variogram(mean_val~1, data = spdf_ger, width = 1, cutoff = 300)
vg_fit_ger <- fit.variogram(vg_ger, vgm("Exp")) 

# Kriging
krig_ger <- krige(mean_val~1, 
               locations = spdf_ger, 
               newdata = sp_grid_ger_sf, 
               model = vg_fit_ger)
```

```{r state_pred_ger}
# Get point from output above
max_point1 <- mean_pred_ger

check_state_ger <- function(max_point) {
  max_point2 <- data.frame(
    longitude = c(max_point$lon),
    latitude = c(max_point$lat)) 
  point_sf <- st_as_sf(max_point2, coords = c("longitude", "latitude"), crs = st_crs(ger_states_plot))
  
  # Check if the point is within any state
  point_within_state <- st_contains(ger_states_plot, point_sf, sparse = FALSE)
  
  # Check which state the point is within (if any)
  state_index <- which(point_within_state, arr.ind = TRUE)
  state_in <- ifelse(length(state_index) > 0, ger_states$name[state_index[1]], "Outside any state")
  
  print(state_in)
}

# run it
state_result <- check_state_ger(max_point1)
print(state_result)
```

```{r ger_map}
legend_max_ger <- round(max(krig_ger$var1.pred), digits = 2)
legend_min_ger <- round(min(krig_ger$var1.pred), digits = 2) 

ggplot() +
    geom_sf(data = krig_ger, aes(fill=var1.pred), shape = 21, size = 0.75, stroke = 0, lwd = 0) +
    geom_sf(data = ger_plot, aes(geometry = geometry), color="black", fill=NA, size = 0.5) +
    geom_sf_text(data = cities_ger_sf, aes(label = City), size = 2.5, nudge_x = 0, nudge_y = -0.15, family = "Optima") +
    geom_sf(data = cities_ger_sf, aes(geometry = geometry), shape = 4) +
    theme_minimal() +
    scale_fill_steps(low="#ffffff", high="#310d59", n.breaks = 20, labels = NULL) +
    theme(axis.title.x = element_blank(), 
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          panel.grid.major = element_blank()) +
    theme(legend.position = c(0.89, 0.35), 
          legend.title = element_text(size = 6, family = "Optima"), 
          legend.text = element_text(size = 6, family = "Optima", hjust = 1),
          legend.title.align = 0,
          legend.key.size = unit(0.3, "cm"),
          legend.key.width = unit(0.4,"cm")) +
    ggplot2::annotate(geom="text", x = 14.9, y = 50.3, label=legend_max_ger, size = 1.5, family = "Optima") +
    ggplot2::annotate(geom="text", x = 14.9, y = 49.3, label=legend_min_ger, size = 1.5, family = "Optima") +
    labs(fill = " ")

# jpeg("adios3.jpg", units="in", width=3.78, height=4.8, res=300)
```

```{r ger_max}
mean_pred_ger <- calculation_ger %>% 
  filter(mean_val > 0) %>% 
  arrange(desc(mean_val)) %>%
  slice (1:1) %>% 
  dplyr::select(mean_val, lon, lat)
  
median_pred_ger <- calculation_ger %>% 
  filter(median_val > 0) %>% 
  arrange(desc(median_val)) %>%
  slice (1:1) %>% 
  dplyr::select(median_val, lon, lat) 
```

## Austria

```{r aus_kriging}
# get data for strings
kriged_values_aus <- kriged_aus %>% select(c("lon", "lat"))
for (cname in strings) {
  if (cname %in% colnames(kriged_aus)) {
    kriged_values_aus[, paste0(cname)] <- as.numeric(unlist(kriged_aus[, cname]))
  }
}

# create new list of column names for further operations
strings <- toString(colnames(kriged_values_aus[3:ncol(kriged_values_aus)]))
strings <- trimws(unlist(strsplit(strings, ",")))

data_for_calculations_aus <- kriged_values_aus %>% select(-lon, -lat)
calculation_aus <- kriged_values_aus %>% select(lon, lat)
calculation_aus$mean_val <- rowMeans(data_for_calculations_aus)
calculation_aus$median_val <- apply(data_for_calculations_aus, 1, median)
calculation_aus$sd_val <- apply(data_for_calculations_aus, 1, sd)
  
calculation_aus_sf <- st_as_sf(calculation_aus, coords = c("lon", "lat"), crs = crs2)

# get max
mean_pred_aus <- calculation_aus %>% 
  filter(mean_val > 0) %>% 
  arrange(desc(mean_val)) %>%
  slice (1:1) %>% 
  dplyr::select(mean_val, lon, lat)

median_pred_aus <- calculation_aus %>% 
  filter(median_val > 0) %>% 
  arrange(desc(median_val)) %>%
  slice (1:1) %>% 
  dplyr::select(median_val, lon, lat)

# Variogram
data_aus <- calculation_aus["mean_val"]
coords_aus <- calculation_aus[c("lon", "lat")]
spdf_aus <- SpatialPointsDataFrame(coords      = coords_aus,
                               data        = data_aus, 
                               proj4string = crs3)
spdf_aus = spdf_aus[which(!duplicated(spdf_aus@coords)), ]
vg_aus <- variogram(mean_val~1, data = spdf_aus, width = 1, cutoff = 300)
vg_fit_aus <- fit.variogram(vg_aus, vgm("Exp")) 

# Kriging
krig_aus <- krige(mean_val~1, 
               locations = spdf_aus, 
               newdata = sp_grid_aus_sf, 
               model = vg_fit_aus)
```

```{r state_pred_aus}
# Get point from output above
max_point1 <- mean_pred_aus

check_state_aus <- function(max_point) {
  max_point2 <- data.frame(
    longitude = c(max_point$lon),
    latitude = c(max_point$lat)) 
  point_sf <- st_as_sf(max_point2, coords = c("longitude", "latitude"), crs = st_crs(aus_states_plot))
  
  # Check if the point is within any state
  point_within_state <- st_contains(aus_states_plot, point_sf, sparse = FALSE)
  
  # Check which state the point is within (if any)
  state_index <- which(point_within_state, arr.ind = TRUE)
  state_in <- ifelse(length(state_index) > 0, aus_states$name[state_index[1]], "Outside any state")
  
  print(state_in)
}

# run it
state_result <- check_state_aus(max_point1)
print(state_result)
```


```{r aus_map}
legend_max_aus <- round(max(krig_aus$var1.pred), digits = 2)
legend_min_aus <- round(min(krig_aus$var1.pred), digits = 2) 

ggplot() +
    geom_sf(data = krig_aus, aes(fill=var1.pred), shape = 21, size = 0.75, stroke = 0, lwd = 0) +
    geom_sf(data = aus_plot, aes(geometry = geometry), color="black", fill=NA, size = 0.5) +
    geom_sf_text(data = cities_aus_sf, aes(label = City), size = 2.5, nudge_x = 0, nudge_y = -0.15, family = "Optima") +
    geom_sf(data = cities_aus_sf, aes(geometry = geometry), shape = 4) +
    theme_minimal() +
    scale_fill_steps(low="#ffffff", high="#310d59", n.breaks = 20, labels = NULL) +
    theme(axis.title.x = element_blank(), 
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          panel.grid.major = element_blank()) +
    theme(legend.position = c(0.12, 0.85), 
          legend.title = element_text(size = 8, family = "Optima"), 
          legend.text = element_text(size = 6, family = "Optima", hjust = 1),
          legend.title.align = 0,
          legend.key.size = unit(0.3, "cm"),
          legend.key.width = unit(0.4,"cm")) +
  ggplot2::annotate(geom="text", x = 10.6, y = 49.05, label=legend_max_aus, size = 1.5, family = "Optima") +
  ggplot2::annotate(geom="text", x = 10.6, y = 48.2, label=legend_min_aus, size = 1.5, family = "Optima") +
    labs(fill = " ")

#tiff("adios.jpg", units="in", width=3.5, height=2, res=300)
```

```{r aus_max}
mean_pred_aus <- calculation_aus %>% 
  filter(mean_val > 0) %>% 
  arrange(desc(mean_val)) %>%
  slice (1:1) %>% 
  dplyr::select(mean_val, lon, lat)
  
median_pred_aus <- calculation_aus %>% 
  filter(median_val > 0) %>% 
  arrange(desc(median_val)) %>%
  slice (1:1) %>% 
  dplyr::select(median_val, lon, lat) 
```

## Switzerland

```{r swi_kriging}
# get data for strings
kriged_values_swi <- kriged_swi %>% select(c("lon", "lat"))

for (cname in strings) {
  if (cname %in% colnames(kriged_swi)) {
    kriged_values_swi[, paste0(cname)] <- as.numeric(unlist(kriged_swi[, cname]))
  }
}

# create new list of column names for further operations
strings <- toString(colnames(kriged_values_swi[3:ncol(kriged_values_swi)]))
strings <- trimws(unlist(strsplit(strings, ",")))

data_for_calculations_swi <- kriged_values_swi %>% select(-lon, -lat)
calculation_swi <- kriged_values_swi %>% select(lon, lat)
calculation_swi$mean_val <- rowMeans(data_for_calculations_swi)
calculation_swi$median_val <- apply(data_for_calculations_swi, 1, median)
calculation_swi$sd_val <- apply(data_for_calculations_swi, 1, sd)
  
calculation_swi <- data.frame(calculation_swi, sign_mean = sign(calculation_swi$mean_val))
calculation_swi <- data.frame(calculation_swi, sign_median = sign(calculation_swi$median_val))

calculation_swi_sf <- st_as_sf(calculation_swi, coords = c("lon", "lat"), crs = crs2)

# get max
mean_pred_swi <- calculation_swi %>% 
  filter(mean_val > 0) %>% 
  arrange(desc(mean_val)) %>%
  slice (1:1) %>% 
  dplyr::select(mean_val, lon, lat)
  
median_pred_swi <- calculation_swi %>% 
  filter(median_val > 0) %>% 
  arrange(desc(median_val)) %>%
  slice (1:1) %>% 
  dplyr::select(median_val, lon, lat) 

# Variogram
data_swi <- calculation_swi["mean_val"]
coords_swi <- calculation_swi[c("lon", "lat")]
spdf_swi <- SpatialPointsDataFrame(coords      = coords_swi,
                               data        = data_swi, 
                               proj4string = crs3)
spdf_swi = spdf_swi[which(!duplicated(spdf_swi@coords)), ]
vg_swi <- variogram(mean_val~1, data = spdf_swi, width = 1, cutoff = 300)
vg_fit_swi <- fit.variogram(vg_swi, vgm("Exp")) 

# Kriging
krig_swi <- krige(mean_val~1, 
               locations = spdf_swi, 
               newdata = sp_grid_swi_sf, 
               model = vg_fit_swi)
```

```{r state_pred_swi}
# Get point from output above
max_point1 <- mean_pred_swi

check_state_swi <- function(max_point) {
  max_point2 <- data.frame(
    longitude = c(max_point$lon),
    latitude = c(max_point$lat)) 
  point_sf <- st_as_sf(max_point2, coords = c("longitude", "latitude"), crs = st_crs(swi_states_plot))
  
  # Check if the point is within any state
  point_within_state <- st_contains(swi_states_plot, point_sf, sparse = FALSE)
  
  # Check which state the point is within (if any)
  state_index <- which(point_within_state, arr.ind = TRUE)
  state_in <- ifelse(length(state_index) > 0, swi_states$name[state_index[1]], "Outside any state")
  
  print(state_in)
}

# run it
state_result <- check_state_swi(max_point1)
print(state_result)
```

```{r swi_map}
legend_max_swi <- round(max(krig_swi$var1.pred), digits = 2)
legend_min_swi <- round(min(krig_swi$var1.pred), digits = 2) 

ggplot() +
    geom_sf(data = krig_swi, aes(fill=var1.pred), shape = 21, size = 0.75, stroke = 0, lwd = 0) +
    geom_sf(data = swi_plot, aes(geometry = geometry), color="black", fill=NA, size = 0.5) +
    geom_sf_text(data = cities_swi_sf, aes(label = City), size = 2.5, nudge_x = 0, nudge_y = -0.15, family = "Optima") +
    geom_sf(data = cities_swi_sf, aes(geometry = geometry), shape = 4) +
    theme_minimal() +
    scale_fill_steps(low="#ffffff", high="#310d59", n.breaks = 20, labels = NULL) +
    theme(axis.title.x = element_blank(), 
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          panel.grid.major = element_blank()) +
    theme(legend.position = c(-0.02, 0.8), 
          legend.title = element_text(size = 6, family = "Optima"), 
          legend.text = element_text(size = 6, family = "Optima", hjust = 1),
          legend.title.align = 0,
          legend.key.size = unit(0.3, "cm"),
          legend.key.width = unit(0.4,"cm")) +
    ggplot2::annotate(geom="text", x = 5.95, y = 47.7, label=legend_max_swi, size = 1.5, family = "Optima") +
    ggplot2::annotate(geom="text", x = 5.95, y = 47.1, label=legend_min_swi, size = 1.5, family = "Optima") +
    labs(fill = " ")

# jpeg("adios3.jpg", units="in", width=3.5, height=2, res=300)
```

```{r swi_max}
mean_pred_swi <- calculation_swi %>% 
  filter(mean_val > 0) %>% 
  arrange(desc(mean_val)) %>%
  slice (1:1) %>% 
  dplyr::select(mean_val, lon, lat)
  
median_pred_swi <- calculation_swi %>% 
  filter(median_val > 0) %>% 
  arrange(desc(median_val)) %>%
  slice (1:1) %>% 
  dplyr::select(median_val, lon, lat) 
```










### some old code as backup

```{r old}
  #calculation <- data.frame(calculation, sign_mean = sign(calculation$mean_val))
  #calculation <- data.frame(calculation, sign_median = sign(calculation$median_val))

    #kriged_data <- merge(x = kriged_data, y = krig3[ , 3], by.x = "geometry", by.y = "geometry", all.x=TRUE)
    #kriged_data <- st_join(kriged_data, krig3)
    #https://r-spatial.github.io/sf/reference/st_join.html
    #combined <- kriged_data %>% st_join(krig3)
#krig3_notsf <- dplyr::select(krig3_notsf, qword, "x", "y")

write.csv(calculation,"log_ransom_calculation.csv", row.names = FALSE)
write.csv(morans_calc,"new_ransom_all_morans.csv", row.names = FALSE)

calculation_ger <- data.frame(calculation_ger, sign_mean = sign(calculation_ger$mean_val))
calculation_ger <- data.frame(calculation_ger, sign_median = sign(calculation_ger$median_val))
```