---
title: "Dialect region shapefiles"
author: "Dana Roemling"
date: "2024-10-18"
output: html_document
toc: true
toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Dialect Region Shapefiles

In order to check if a hotspot falls into a specific dialect area, this rmd documents how to set up the dialect shapefiles and then uses them for hotspot checking.

Originally, the data for this shapefile comes from the Rede project, Regionalsprache.de. They provide a mapping tool for a lot of different dialect datasets. They have a raster version of the dialect regions found by Lameli. It can be accessed here: https://www.regionalsprache.de/SprachGis/RasterMap/Dialekteinteilungen/@32314

The corresponding academic publication for Lameli is:
Lameli, Alfred (2013): Strukturen im Sprachraum: Analysen zur arealtypologischen Komplexität der Dialekte in Deutschland. Berlin & New York: De Gruyter.

In order to use the raster data, it needed to be converted into a shapefile. The download from Rede offers to export it as a Geojson file, which is the closest file format to what I need. 

I then used https://mapshaper.org which offers to convert spatial file formats. Here I could change the Geojson to an ESRI format, with which I could work further.

```{r data_libs}
# Load the packages
library(sf)       # For handling spatial data
library(sp)       # For setting the CRS
library(dplyr)    # For data manipulation
library(ggplot2)  # For plotting spatial data
library(ggpattern) # For different fill options when plotting
library(paletteer) # For colour palette

# Read the spatial data from a shapefile using the sf package
# 'dsn' specifies the file path to the shapefile directory
# 'layer' specifies the name of the shapefile (without the .shp extension)
# This is the old shapefile
germany <- st_read(dsn = "/Users/dana/Documents/R/PHD/data_maps/dialektregion",
                   layer = "lameli")

# This is the exported shapefile from below, to be read in for further work
germany <- st_read(dsn = "/Users/dana/Documents/R/PHD/data_maps/dialektregion",
                   layer = "dialect")
colnames(germany) <- c("gid", "region", "transition", "lowsaxon", 
                       "highgerman", "ripuarian", "geometry")

# Transform into four broader dialect areas
germany$language_region <- ifelse(germany$ripuarian == "ja", "Ripuarian",
                           ifelse(germany$lowsaxon == "ja", "Low German",
                           ifelse(germany$highgerman == "ja", "High German", "Middle German")))

# Transform the coordinate reference system (CRS) of the 'germany' object
# This converts the spatial data to the WGS84 coordinate system, which uses latitude and longitude
# '+proj=longlat' specifies the use of geographic coordinates (latitude/longitude)
# '+datum=WGS84' ensures the datum is the World Geodetic System 1984, a standard for global positioning
# 'no_defs' avoids loading other CRS parameters from external sources
# '+ellps=WGS84' specifies the shape of the Earth (ellipsoid) as WGS84
# '+towgs84=0,0,0' applies no transformation offsets since it's already in WGS84
germany <- st_transform(germany, 
                        CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"))

# Create a new object 'germany_small' that is a subset of the 'germany' object
germany_small <- germany %>% select(gid, geometry)
```

To make the shapefile useful, I need to add some information to it. 
This was done using https://mapshaper.org, which has a select tool to see the IDs of the polygons. These were then compared to Lameli (2013) to name all regions. 
In case Lameli (2013) did not name the regions appropriately, they were double-checked with this publication:
Jürgen Erich Schmidt. Vom traditionellen Dialekt zu den modernen deutschen Regionalsprachen. In Deutsche Akademie für Sprache und Dichtung / Union der deutschen Akademien der Wissenschaften (Hrsg.), Vielfalt und Einheit der deutschen Sprache: Zweiter Bericht zur Lage der deutschen Sprache, 105-143. Tübingen: Stauffenburg.
Accessible as an image on Rede: https://www.regionalsprache.de/SprachGis/VectorMap/Dialekteinteilungen/@54860

```{r data_wrangling1}
# Create a new column called 'region' in the 'germany_small' data frame
# Initialize all values in this column to NA (missing values)
# This column is intended to store regional labels for each spatial unit
germany_small$region <- NA

# Create another new column called 'übergang' in the 'germany_small' data frame
# Initialize all values in this column to the string "nein" (German for "no")
# This column is a categorical variable indicating whether or not a region is a transition area
germany_small$übergang <- "nein"

# Create a new column called 'niederdeutsch' in the 'germany_small' data frame
# Initialize all values in this column to NA (missing values)
# This column is storing information about whether the region is classified as speaking Niederdeutsch (Low German)
# Same for the High German and ripuarisch dialect areas
germany_small$niederdeutsch <- NA
germany_small$hochdeutsch <- NA
germany_small$ripuarisch <- NA
```

```{r data_wrangling2}
# Example annotation
# Assign the value "nordostniederdeutsch" to the first element of the 'region' column in the 'germany_small' data frame
# This means that the first geographic unit (row 1) in 'germany_small' is classified as belonging to the "nordostniederdeutsch" region
germany_small$region[1] <- "nordostniederdeutsch"

# The rest of the batch
germany_small$region[2] <- "nordniederdeutsch"
germany_small$region[4] <- "nordniederdeutsch"
germany_small$region[25] <- "nordniederdeutsch"
germany_small$region[14] <- "westniederdeutsch"
germany_small$region[24] <- "brandenburgisch"
germany_small$region[6] <- "brandenburgisch"
germany_small$region[26] <- "ostfälisch"
germany_small$region[27] <- "ostfälisch"
germany_small$region[21] <- "westfälisch"
germany_small$region[22] <- "westfälisch"
germany_small$region[23] <- "westfälisch"
germany_small$region[20] <- "westfälisch"
germany_small$region[13] <- "niederfränkisch"
germany_small$region[3] <- "nordwestniederdeutsch"
germany_small$region[9] <- "bairisch"
germany_small$region[7] <- "ostfränkisch"
germany_small$region[16] <- "ostfränkisch"
germany_small$region[5] <- "schwäbisch"
germany_small$region[12] <- "ripuarisch"
germany_small$region[11] <- "moselfränkisch"
germany_small$region[10] <- "rheinfränkisch-hessisch"
germany_small$region[19] <- "rheinfränkisch-hessisch"
germany_small$region[8] <- "sächsisch-thüringisch"
germany_small$region[17] <- "ostfränkisch"
germany_small$region[15] <- "schwäbisch"
germany_small$region[18] <- "schwäbisch"
germany_small$region[28] <- "sächsisch-thüringisch"

# Annotation for the transition areas
germany_small$übergang[24] <- "ja"
germany_small$übergang[25] <- "ja"
germany_small$übergang[4] <- "ja"
germany_small$übergang[27] <- "ja"
germany_small$übergang[22] <- "ja"
germany_small$übergang[23] <- "ja"
germany_small$übergang[20] <- "ja"
germany_small$übergang[13] <- "ja"
germany_small$übergang[11] <- "ja"
germany_small$übergang[17] <- "ja"
germany_small$übergang[15] <- "ja"
germany_small$übergang[18] <- "ja"
germany_small$übergang[16] <- "ja"
germany_small$übergang[19] <- "ja"
germany_small$übergang[28] <- "ja"

# Annotation for Low German dialect area
germany_small$niederdeutsch[1] <- "ja"
germany_small$niederdeutsch[2] <- "ja"
germany_small$niederdeutsch[4] <- "ja"
germany_small$niederdeutsch[25] <- "ja"
germany_small$niederdeutsch[14] <- "ja"
germany_small$niederdeutsch[24] <- "ja"
germany_small$niederdeutsch[6] <- "ja"
germany_small$niederdeutsch[26] <- "ja"
germany_small$niederdeutsch[27] <- "ja"
germany_small$niederdeutsch[21] <- "ja"
germany_small$niederdeutsch[22] <- "ja"
germany_small$niederdeutsch[23] <- "ja"
germany_small$niederdeutsch[20] <- "ja"
germany_small$niederdeutsch[3] <- "ja"

# Annotation for High German dialect area
germany_small$hochdeutsch[9] <- "ja"
germany_small$hochdeutsch[5] <- "ja"
germany_small$hochdeutsch[15] <- "ja"
germany_small$hochdeutsch[16] <- "ja"
germany_small$hochdeutsch[7] <- "ja"
germany_small$hochdeutsch[17] <- "ja"

# Annotation for ripuarisch dialect area
germany_small$ripuarisch[12] <- "ja"
germany_small$ripuarisch[11] <- "ja"
germany_small$ripuarisch[13] <- "ja"

# code the rest as no
germany_small[is.na(germany_small)] <- "nein"
```

```{r map_export}
# Create a ggplot object to visualize the data
# ggplot() initializes a blank plot, providing a base for adding layers
ggplot() +
  # Add a layer for plotting simple features (spatial data) using the geom_sf() function
  # 'data = germany_small' specifies that the spatial data to be plotted comes from the 'germany_small' data frame
  # aes(geometry = geometry) maps the geometry column (which contains spatial shapes) to the plot
  # This will display the geographical shapes from the 'germany_small' data frame on the plot
  geom_sf(data = germany_small, 
          aes(geometry = geometry))

# Write the 'germany_small' data frame to a shapefile using the st_write() function from the sf package
# The shapefile will be saved with the name "dialect_shp.shp" in the current working directory
# This operation exports the spatial data, preserving its structure and attributes, making it available for use in GIS software or other spatial analysis tools
st_write(germany_small, "dialect.shp")
```

```{r id_map}
# Fix invalid geometries
# This is necessary because some of the borders overlap, 
# which can lead to issues when performing spatial operations.
germany <- st_transform(germany, 
                        crs = 32632)  
# Transform the geometries to UTM zone 32N (meters) for valid geometric operations
germany <- st_make_valid(germany)  # Fix any invalid geometries in the dataset
germany <- st_transform(germany, 
                        crs = 4326)  
# Transform back to WGS84 (latitude/longitude) for consistency

# Create unique dialect IDs based on language_region
germany <- germany %>%
  # Calculate centroids for each geometry to help with ordering
  mutate(centroid = st_centroid(geometry)) %>%
  # Extract the longitude of each centroid for ordering
  mutate(lon = st_coordinates(centroid)[, 1]) %>%
  # Arrange the data frame by longitude for west-to-east ordering
  arrange(lon) %>%
  # Group by language_region to ensure IDs are created within each region
  group_by(language_region) %>%
  # Create a new dialect_id by concatenating the first letter of 
  # language_region with a unique number
  mutate(dialect_id = paste0(toupper(substring(language_region, 
                                               1, 
                                               1)), 
                             row_number())) %>%
  ungroup() %>%
  # Remove temporary columns used for calculations
  select(-centroid, -lon)  # Clean up by removing temporary columns

# Plot the dialect regions with updated IDs
ggplot() +
  # Draw the geometries for the dialect regions, filling with white
  geom_sf(data = germany, 
          aes(geometry = geometry),
          fill = "white") +
  # Add text labels with the dialect IDs to the map
  geom_sf_text(data = germany, 
               aes(label = dialect_id), 
               size = 3, # Adjust the size of the text labels as necessary
               position = position_nudge(y = 0.02), 
               # Adjust position to avoid overlap with other text
               check_overlap = TRUE) + # Prevent text labels from overlapping on the map
  theme_minimal() +  # Use a minimal theme for the plot
  theme(axis.title.x = element_blank(),  # Remove x-axis title
        axis.title.y = element_blank(),  # Remove y-axis title
        axis.text.x = element_blank(),    # Remove x-axis text labels
        axis.text.y = element_blank(),    # Remove y-axis text labels
        panel.grid.major = element_blank(), # Remove major grid lines
        plot.title = element_text(size = 10, family = "Optima", hjust = 0.5) 
        # Customize the title's appearance
        ) +
  # Set the title of the plot
  ggtitle("Dialect Regions within Germany")
```

```{r publish_plot}
# plot with just dialect regions
ggplot() +
  geom_sf(data = germany, 
          aes(geometry = geometry),
          fill = "white") +
  theme_minimal() +
  theme(axis.title.x = element_blank(), 
        axis.title.y = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        panel.grid.major = element_blank(),
        plot.title = element_text(size = 10, family = "Optima", hjust = 0.5)
        ) +
  ggtitle("Dialect Regions within Germany")
```

```{r publish_plot2}
# plot with transition areas patterned
ggplot(germany) + 
  geom_sf(data = germany, 
          aes(geometry = geometry, 
              fill = transition), 
          color = 'black',
          fill = "white") +  # Basic fill for all areas
  geom_sf_pattern(data = subset(germany, 
                                transition == "ja"),  # Apply pattern only to ja areas
                  aes(geometry = geometry),
                  pattern = 'stripe',
                  fill = 'white',  # Base fill for pattern areas
                  pattern_density = 0.04,
                  pattern_spacing = 0.02,
                  pattern_fill = 'black',
                  color = 'black',
                  pattern_colour  = 'black'
  ) +
  scale_pattern_fill_manual(values = c(ja = 'black')) +  
  theme_minimal() +
  theme(axis.title.x = element_blank(), 
        axis.title.y = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        panel.grid.major = element_blank(),
        plot.title = element_text(size = 10, family = "Optima", hjust = 0.5),
        legend.position = "none"
  ) +
  ggtitle("Transition Dialect Areas")
```

```{r publish_plot3}
# plot with low saxon coloured
ggplot() +
  geom_sf(data = germany, 
          aes(geometry = geometry,
              fill = lowsaxon),
          alpha = 0.3) +
  scale_fill_manual(values = c("plum4", "white")) +
  theme_minimal() +
  theme(axis.title.x = element_blank(), 
        axis.title.y = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        panel.grid.major = element_blank(),
        plot.title = element_text(size = 10, family = "Optima", hjust = 0.5),
        legend.position = "none"
  ) +
  ggtitle("Low vs. Middle and High German")
```

```{r publish_plot4}
# palette uses paletteer
# https://r-graph-gallery.com/color-palette-finder

# four major areas coloured
ggplot() +
  geom_sf(data = germany, 
          aes(geometry = geometry,
              fill = language_region),
          alpha = 0.3) +
  scale_fill_paletteer_d("MoMAColors::Fritsch") +
  theme_minimal() +
  theme(axis.title.x = element_blank(), 
        axis.title.y = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        panel.grid.major = element_blank(),
        plot.title = element_text(size = 10, family = "Optima", hjust = 0.5),
        legend.position = "none"
  ) +
  ggtitle("Four Major Dialect Regions")
```

## Hotspot Prediction

Following from rmd 16, these regions can be used to extract in which dialect area the hotspot of a map falls.

```{r hotspot_prep}
# Load necessary libraries
library(dplyr)          # For data manipulation and pipeline (%>%) operations
library(spdep)          # For spatial analysis
library(sf)             # For handling spatial data using simple features
library(sp)             # For handling spatial data, particularly the older sp objects
library(tidyverse)      # Includes multiple packages like ggplot2, dplyr, tidyr
library(gstat)          # For geostatistical modeling and kriging
library(stringr)        # For string manipulation
library(rnaturalearth)  # For downloading and using natural Earth map data

# Load kriging data for Germany
kriged_ger <- read.csv(file = 'data_ling/kriged/kriged_5k_300cutoff_germany.csv', 
                       header = TRUE)  
# Read the CSV file containing kriging results into a data frame

# Clean up column names by removing dots and "X" prefixes from column names
names(kriged_ger) <- gsub("\\.", "", names(kriged_ger))  # Remove all dots from the column names
names(kriged_ger) <- gsub("\\<X", "", names(kriged_ger)) # Remove "X" prefix from the column names

# Select the first 2002 columns of the kriging data (assuming they are relevant for further analysis)
kriged_ger <- kriged_ger %>% dplyr::select(1:2002)

# Get the outline of Germany using Natural Earth data with high resolution
ger_outline <- ne_countries(country = "Germany", 
                            returnclass = "sf", 
                            scale = "large")  # Load Germany's map as an sf object

# Extract only the geometry of Germany for plotting purposes
ger_plot <- ger_outline %>% 
  dplyr::select(geometry)  # Select only the geometry column from the sf object

# Convert the simple features object to a Spatial object 
ger_spatial <- as_Spatial(ger_plot)  # Convert sf object to a Spatial object using as_Spatial()

# Create a data frame of major German cities with their coordinates
cities_ger <- data.frame(
  City = c("Köln", "München", "Berlin", "Hamburg", "Frankfurt", "Kassel"),  # List of cities
  Long = c(6.9578, 11.5755, 13.3833, 10, 8.6821, 9.4797),                   # Longitudes 
  Lat = c(50.9422, 48.1372, 52.5167, 53.55, 50.1109, 51.3127))              # Latitudes 

# Set coordinate reference systems
crs2 <- st_crs(4326)
crs3 <- CRS("+init=epsg:4326") 

# Convert the cities data frame to a simple features object (sf), 
# using the specified coordinate columns
cities_ger_sf <- st_as_sf(cities_ger, coords = c("Long", "Lat"), crs = crs2)  
# Create sf object with coordinates and CRS

# Create a systematic grid of points over the spatial extent of Germany for spatial sampling
sp_grid_ger <- as.data.frame(spsample(ger_spatial,
                                      n = 100000,           
                                      # Number of points to sample in the grid
                                      type = "regular",     
                                      # Systematic sampling in a regular grid pattern
                                      offset = c(0.5, 0.5))) 
                                      # Ensures grid alignment is consistent every time

# Convert the regular grid points into a simple features (sf) object for spatial operations
sp_grid_ger_sf <- st_as_sf(sp_grid_ger, 
                           coords = c("x1", "x2"), 
                           crs = crs2)  # Convert grid points to sf object

# Create a list of German words for analysis
strings <- tolower(c("aber", "eigentlich", "isses", "mir", "echt", "egal"))  
                    # Convert words to lowercase 

# Ensure that the list of words contains unique values only (removes duplicates)
strings <- unique(strings)  # Remove any duplicate words from the list
```


```{r hotspot_prep2}
# Extract relevant columns (longitude and latitude) from the kriged_ger data frame
kriged_values_ger <- kriged_ger %>% 
  select(c("lon", 
           "lat"))

# Loop through each string in the predefined list of strings
for (cname in strings) {
  # Check if the current string exists as a column in the kriged_ger data frame
  if (cname %in% colnames(kriged_ger)) {
    # If it exists, convert the column to numeric and add it to kriged_values_ger
    kriged_values_ger[, paste0(cname)] <- as.numeric(unlist(kriged_ger[, cname]))
  }
}

# Create a new list of column names for further operations, excluding 'lon' and 'lat'
strings <- toString(colnames(kriged_values_ger[3:ncol(kriged_values_ger)])) 
# Convert column names to a single string
strings <- trimws(unlist(strsplit(strings, ","))) 
# Split the string back into a list of names and trim whitespace

# Prepare the data for calculations by removing the 'lon' and 'lat' columns
data_for_calculations_ger <- kriged_values_ger %>% 
  select(-lon, -lat)

# Create a new data frame for calculations containing only the coordinates
calculation_ger <- kriged_values_ger %>% 
  select(lon, lat)

# Calculate the mean value for each row across the selected columns
calculation_ger$mean_val <- rowMeans(data_for_calculations_ger)

# Calculate the median value for each row across the selected columns
calculation_ger$median_val <- apply(data_for_calculations_ger, 
                                    1, 
                                    median)

# Calculate the standard deviation for each row across the selected columns
calculation_ger$sd_val <- apply(data_for_calculations_ger, 
                                1, 
                                sd)

# Convert the calculation data frame into a spatial data frame with specified coordinates and coordinate reference system
calculation_ger_sf <- st_as_sf(calculation_ger, 
                               coords = c("lon", "lat"), 
                               crs = crs2)

# Retrieve the maximum mean value, filtering for mean values greater than 0
mean_pred_ger <- calculation_ger %>% 
  filter(mean_val > 0) %>%           # Keep only rows where mean_val is greater than 0
  arrange(desc(mean_val)) %>%        # Sort the data frame in descending order by mean_val
  slice(1:1) %>%                     # Select the top row (highest mean value)
  dplyr::select(mean_val, lon, lat)  # Select only the relevant columns for further analysis

# Not needed for point prediction; perform variogram analysis
data_ger <- calculation_ger["mean_val"]  # Prepare data frame with mean_val for variogram
coords_ger <- calculation_ger[c("lon", "lat")]  # Prepare data frame with coordinates
# Create a spatial points data frame with the coordinates and data
spdf_ger <- SpatialPointsDataFrame(coords = coords_ger,
                                    data = data_ger, 
                                    proj4string = crs3)

# Remove duplicate coordinates to avoid issues in variogram calculation
spdf_ger = spdf_ger[which(!duplicated(spdf_ger@coords)), ]

# Compute the empirical variogram for the mean value
vg_ger <- variogram(mean_val~1, 
                    data = spdf_ger, 
                    width = 1, 
                    cutoff = 300)

# Fit a variogram model to the empirical variogram
vg_fit_ger <- fit.variogram(vg_ger, 
                            vgm("Exp")) 

# Perform Kriging interpolation using the fitted variogram model
krig_ger <- krige(mean_val~1, 
                  locations = spdf_ger, 
                  newdata = sp_grid_ger_sf, 
                  model = vg_fit_ger)  # Use the variogram model for prediction
```

```{r prediction_region}
# Get point from output above
max_point1 <- mean_pred_ger

# Fix invalid geometries
# This is because some of the borders overlap 
germany <- st_transform(germany, 
                        crs = 32632)
germany <- st_make_valid(germany)
germany <- st_transform(germany, 
                        crs = 4326)

# Function to check which dialect region the point falls into
check_dialect_region <- function(max_point) {
  # Create a data frame for the highest point's coordinates
  max_point_df <- data.frame(
    longitude = c(max_point$lon),
    latitude = c(max_point$lat)
  )
  
  # Convert the point to an sf object with the same CRS as the dialect regions
  point_sf <- st_as_sf(max_point_df, 
                       coords = c("longitude", "latitude"), 
                       crs = st_crs(germany))
  
  # Ensure that the point's CRS matches the germany object
  point_sf <- st_transform(point_sf, 
                           st_crs(germany))

  # Check if the point is within any dialect region
  point_within_region <- st_contains(germany, 
                                     point_sf, 
                                     sparse = FALSE)
  
  # Find which dialect region the point is within (if any)
  region_index <- which(point_within_region, 
                        arr.ind = TRUE)
  
  # Initialize a list to store the results
  result <- list()
  
  # If the point is within a region, extract the relevant information
  if (length(region_index) > 0) {
    result$region_name <- germany$region[region_index[1]]
    result$transition <- germany$transition[region_index[1]]
    result$language_region <- germany$language_region[region_index[1]]
    result$gid <- germany$gid[region_index[1]]
  } else {
    result$region_name <- "Outside any region"
    result$transition <- NA
    result$language_region <- NA
    result$gid <- NA
  }
  
  # Print the result
  print(result)
  
  return(result)  # Return the result for further use
}

# Run the function to check which dialect region the highest point lies in
region_result <- check_dialect_region(max_point1)
print(region_result)
```






