---
title: "Grids Over Germany"
author: "Dana Roemling"
date: "2024-10-21"
output: html_document
toc: true
toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Grids for Evaluation

In order to evaluate how well the algorithm is performing, I want to overlay Germany with a grid that has gradually decreasing cell sizes to see how accuracy for prediction changes. 

```{r libraries_setup}
# Spatial Data Libraries
library(sf)          # Modern package for handling vector geospatial data (points, lines, polygons).
library(sp)          # Older spatial data package, still used in some legacy code.
library(rworldmap)   # Provides world map data for plotting base maps.

# Data Manipulation and Visualization
library(tidyverse)   
# Collection of data manipulation and visualization packages (includes ggplot2, dplyr, etc.).
library(ggplot2)     
# Visualization package for creating plots, particularly useful for geospatial data.
library(dplyr)       # Part of the tidyverse, used for efficient data manipulation.

# Visualization Enhancements
library(RColorBrewer) # Creates color palettes for visualizations that are colorblind-friendly.
library(scales)       # Helps format data in visualizations (e.g., axes, labels).

# Load in the Europe shapefile
EU <- st_read(dsn = "data_maps/EU/Try/", 
              layer = "NUTS_RG_20M_2021_3035")

# Change projection so it's all uniform
EU <- st_transform(EU, 
                   CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"))

# Extract only Germany as a whole (not with admin. regions)
# And store in new object
ger <- EU %>% filter(NUTS_ID == "DE")
```

# 4 Cell Grid

```{r 4cells}
# Create a grid (fishnet) based on the outline stored in the 'ger' object
fishnet <- st_make_grid(
  st_transform(ger, crs = st_crs(4326)), # Transform the CRS of 'ger' to WGS84 (longitude/latitude).
  cellsize = 5,       # Set the size of each grid cell to be 5 units.
  square = TRUE,      # Ensure that the grid cells are square-shaped.
  offset = c(0.5, 1)  # Shift the origin of the grid to ensure alignment, starting at longitude 0.5, latitude 1.
) %>% 
  st_sf() # Convert the grid to a simple features (sf) object for spatial operations.

# Intersect the fishnet grid with the 'ger' plot to retain only the grid cells that overlap with the region of interest (GSA region)
grid_ger <- fishnet %>% 
  st_intersection(ger) # Keep only those cells that fall within the 'ger' boundary.

# Assign a unique ID to each cell in the grid for reference
grid_ger$cell_id <- 1:nrow(grid_ger) # Create a column called 'cell_id' with sequential numbers for each cell.

# Save the grid with the cell IDs to a shapefile for future use or analysis
st_write(grid_ger, "4_cell_grid_ger.shp") 
# Write the resulting grid to a shapefile named "4_cell_grid_ger.shp".
```

```{r 4plot}
# Read the shapefile data for the 4-cell grid
grid_4cell <- st_read(
  dsn = "/Users/dana/Documents/R/PHD/data_maps/grids/4cell",  
  # Path to the directory containing the shapefile
  layer = "4_cell_grid_ger"                                   
  # Name of the specific shapefile layer to be read
)

# Calculate centroids of each cell in the grid and add them as a new column in the grid_4cell data
grid_4cell <- grid_4cell %>%
  mutate(centroid = st_centroid(geometry)) 
# Compute the centroid of each cell for text placement later

# Plot the grid and label each cell with its ID
ggplot() +
  geom_sf(data = grid_4cell, 
          aes(geometry = geometry),  # Plot the grid cells using their geometry
          fill = "white") +          # Set the fill color of the cells to white
  geom_sf_text(data = grid_4cell, 
               aes(geometry = centroid, label = cell_id), # Use centroids for label positions
               size = 3,                                 # Set the text size
               hjust = 0.5,                              # Center text horizontally within each cell
               vjust = 0.5                               # Center text vertically within each cell
  ) +
  theme_minimal() +  # Use a minimalistic theme to reduce visual clutter
  theme(
    axis.title.x = element_blank(),       # Remove x-axis title
    axis.title.y = element_blank(),       # Remove y-axis title
    axis.text.x = element_blank(),        # Remove x-axis text (tick labels)
    axis.text.y = element_blank(),        # Remove y-axis text (tick labels)
    panel.grid.major = element_blank(),   # Remove major grid lines
    plot.title = element_text(size = 10, family = "Optima", hjust = 0.5) # Customize title appearance
  ) +
  ggtitle("4 Cell Grid")  # Add a title to the plot
```

## 9 Cell Grid

Annotation same as above, the only changes are made in cellsize and offset. 

```{r 9cell}
fishnet <- st_make_grid(st_transform(ger, 
                                    crs=st_crs(4326)), 
                       cellsize = 3.2,
                       square = TRUE,
                       offset = c(-0.6,-2)
                       ) %>% 
  st_sf()

grid_ger <- fishnet %>% 
  st_intersection(ger)

grid_ger$cell_id <- 1:nrow(grid_ger)

st_write(grid_ger, "9_cell_grid_ger.shp")
```

```{r 9plot}
grid_9cell <- st_read(dsn = "/Users/dana/Documents/R/PHD/data_maps/grids/9cell", 
                      layer = "9_cell_grid_ger")
grid_9cell <- grid_9cell %>%
  mutate(centroid = st_centroid(geometry))

ggplot() +
  geom_sf(data = grid_9cell, 
          aes(geometry = geometry),
          fill = "white") +
  geom_sf_text(data = grid_9cell, 
               aes(geometry = centroid, label = cell_id), 
               size = 3, 
               hjust = 0.5, # Center horizontally
               vjust = 0.5  # Center vertically
  ) +
  theme_minimal() +
  theme(axis.title.x = element_blank(), 
        axis.title.y = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        panel.grid.major = element_blank(),
        plot.title = element_text(size = 10, family = "Optima", hjust = 0.5)
  ) +
  ggtitle("9 Cell Grid")
```

## Max Point Analysis for Regular Grid

As previously, we can check if the hotspot of a word or a message falls into one of the cells of the grid. Similar code in rmd 16 and 20.

```{r hotspot_prep}
# Load necessary libraries
library(dplyr)          # For data manipulation and pipeline (%>%) operations
library(spdep)          # For spatial analysis
library(sf)             # For handling spatial data using simple features
library(sp)             # For handling spatial data, particularly the older sp objects
library(tidyverse)      # Includes multiple packages like ggplot2, dplyr, tidyr
library(gstat)          # For geostatistical modeling and kriging
library(stringr)        # For string manipulation
library(rnaturalearth)  # For downloading and using natural Earth map data

# Load kriging data for Germany
kriged_ger <- read.csv(file = 'data_ling/kriged/kriged_5k_300cutoff_germany.csv', 
                       header = TRUE)  
# Read the CSV file containing kriging results into a data frame

# Clean up column names by removing dots and "X" prefixes from column names
names(kriged_ger) <- gsub("\\.", "", names(kriged_ger))  # Remove all dots from the column names
names(kriged_ger) <- gsub("\\<X", "", names(kriged_ger)) # Remove "X" prefix from the column names

# Select the first 2002 columns of the kriging data (assuming they are relevant for further analysis)
kriged_ger <- kriged_ger %>% dplyr::select(1:2002)

# Get the outline of Germany using Natural Earth data with high resolution
ger_outline <- ne_countries(country = "Germany", 
                            returnclass = "sf", 
                            scale = "large")  # Load Germany's map as an sf object

# Extract only the geometry of Germany for plotting purposes
ger_plot <- ger_outline %>% 
  dplyr::select(geometry)  # Select only the geometry column from the sf object

# Convert the simple features object to a Spatial object 
ger_spatial <- as_Spatial(ger_plot)  # Convert sf object to a Spatial object using as_Spatial()

# Create a data frame of major German cities with their coordinates
cities_ger <- data.frame(
  City = c("Köln", "München", "Berlin", "Hamburg", "Frankfurt", "Kassel"),  # List of cities
  Long = c(6.9578, 11.5755, 13.3833, 10, 8.6821, 9.4797),                   # Longitudes 
  Lat = c(50.9422, 48.1372, 52.5167, 53.55, 50.1109, 51.3127))              # Latitudes 

# Set coordinate reference systems
crs2 <- st_crs(4326)
crs3 <- CRS("+init=epsg:4326") 

# Convert the cities data frame to a simple features object (sf), 
# using the specified coordinate columns
cities_ger_sf <- st_as_sf(cities_ger, coords = c("Long", "Lat"), crs = crs2)  
# Create sf object with coordinates and CRS

# Create a systematic grid of points over the spatial extent of Germany for spatial sampling
sp_grid_ger <- as.data.frame(spsample(ger_spatial,
                                      n = 100000,           
                                      # Number of points to sample in the grid
                                      type = "regular",     
                                      # Systematic sampling in a regular grid pattern
                                      offset = c(0.5, 0.5))) 
                                      # Ensures grid alignment is consistent every time

# Convert the regular grid points into a simple features (sf) object for spatial operations
sp_grid_ger_sf <- st_as_sf(sp_grid_ger, 
                           coords = c("x1", "x2"), 
                           crs = crs2)  # Convert grid points to sf object

# Create a list of German words for analysis
strings <- tolower(c("aber", "eigentlich", "isses", "mir", "echt", "egal"))  
                    # Convert words to lowercase 

# Ensure that the list of words contains unique values only (removes duplicates)
strings <- unique(strings)  # Remove any duplicate words from the list
```


```{r hotspot_prep2}
# Extract relevant columns (longitude and latitude) from the kriged_ger data frame
kriged_values_ger <- kriged_ger %>% 
  select(c("lon", 
           "lat"))

# Loop through each string in the predefined list of strings
for (cname in strings) {
  # Check if the current string exists as a column in the kriged_ger data frame
  if (cname %in% colnames(kriged_ger)) {
    # If it exists, convert the column to numeric and add it to kriged_values_ger
    kriged_values_ger[, paste0(cname)] <- as.numeric(unlist(kriged_ger[, cname]))
  }
}

# Create a new list of column names for further operations, excluding 'lon' and 'lat'
strings <- toString(colnames(kriged_values_ger[3:ncol(kriged_values_ger)])) 
# Convert column names to a single string
strings <- trimws(unlist(strsplit(strings, ","))) 
# Split the string back into a list of names and trim whitespace

# Prepare the data for calculations by removing the 'lon' and 'lat' columns
data_for_calculations_ger <- kriged_values_ger %>% 
  select(-lon, -lat)

# Create a new data frame for calculations containing only the coordinates
calculation_ger <- kriged_values_ger %>% 
  select(lon, lat)

# Calculate the mean value for each row across the selected columns
calculation_ger$mean_val <- rowMeans(data_for_calculations_ger)

# Calculate the median value for each row across the selected columns
calculation_ger$median_val <- apply(data_for_calculations_ger, 
                                    1, 
                                    median)

# Calculate the standard deviation for each row across the selected columns
calculation_ger$sd_val <- apply(data_for_calculations_ger, 
                                1, 
                                sd)

# Convert the calculation data frame into a spatial data frame with specified coordinates and coordinate reference system
calculation_ger_sf <- st_as_sf(calculation_ger, 
                               coords = c("lon", "lat"), 
                               crs = crs2)

# Retrieve the maximum mean value, filtering for mean values greater than 0
mean_pred_ger <- calculation_ger %>% 
  filter(mean_val > 0) %>%           # Keep only rows where mean_val is greater than 0
  arrange(desc(mean_val)) %>%        # Sort the data frame in descending order by mean_val
  slice(1:1) %>%                     # Select the top row (highest mean value)
  dplyr::select(mean_val, lon, lat)  # Select only the relevant columns for further analysis

# Not needed for point prediction; perform variogram analysis
data_ger <- calculation_ger["mean_val"]  # Prepare data frame with mean_val for variogram
coords_ger <- calculation_ger[c("lon", "lat")]  # Prepare data frame with coordinates
# Create a spatial points data frame with the coordinates and data
spdf_ger <- SpatialPointsDataFrame(coords = coords_ger,
                                    data = data_ger, 
                                    proj4string = crs3)

# Remove duplicate coordinates to avoid issues in variogram calculation
spdf_ger = spdf_ger[which(!duplicated(spdf_ger@coords)), ]

# Compute the empirical variogram for the mean value
vg_ger <- variogram(mean_val~1, 
                    data = spdf_ger, 
                    width = 1, 
                    cutoff = 300)

# Fit a variogram model to the empirical variogram
vg_fit_ger <- fit.variogram(vg_ger, 
                            vgm("Exp")) 

# Perform Kriging interpolation using the fitted variogram model
krig_ger <- krige(mean_val~1, 
                  locations = spdf_ger, 
                  newdata = sp_grid_ger_sf, 
                  model = vg_fit_ger)  # Use the variogram model for prediction
```

This is the actual checking. 

```{r hotspot_check}
# get data from above
max_point1 <- mean_pred_ger

# Function to check if the point is within a cell in the grid
check_grid_region <- function(max_point, grid) {
  # Create a data frame for the highest point's coordinates
  max_point_df <- data.frame(
    longitude = c(max_point$lon),
    latitude = c(max_point$lat)
  )
  
  # Convert the point to an sf object with the same CRS as the grid
  point_sf <- st_as_sf(max_point_df, 
                       coords = c("longitude", "latitude"), 
                       crs = st_crs(grid))
  
  # Ensure that the point's CRS matches the grid object
  point_sf <- st_transform(point_sf, st_crs(grid))
  
  # Check if the point is within any grid cell
  point_within_cell <- st_contains(grid, point_sf, sparse = FALSE)
  
  # Find which grid cell the point is within (if any)
  cell_index <- which(point_within_cell, arr.ind = TRUE)
  
  # Initialize a list to store the results
  result <- list()
  
  # If the point is within a cell, extract the relevant information
  if (length(cell_index) > 0) {
    result$cell_id <- grid$cell_id[cell_index[1]]
    result$other_info <- grid$other_info[cell_index[1]]  # Replace with other relevant columns if needed
  } else {
    result$cell_id <- "Outside any cell"
    result$other_info <- NA
  }
  
  # Print the result
  print(result)
  
  return(result)  # Return the result for further use
}

# Run the function to check which grid cell the highest point lies in
grid_result <- check_grid_region(max_point1, grid)

grid_result <- check_grid_region(max_point1, grid_4cell)
print(grid_result)
```


## Hexagon Grid

Using the same parameters as in the 9 cell grid, changing the cells to hexagons leads the following:

```{r 12hexacell}
fishnet <- st_make_grid(st_transform(ger, 
                                    crs=st_crs(4326)), 
                       cellsize = 3.2,
                       square = FALSE,
                       offset = c(-0.6,-2)
                       ) %>% 
  st_sf()

grid_ger <- fishnet %>% 
  st_intersection(ger)

grid_ger$cell_id <- 1:nrow(grid_ger)

st_write(grid_ger, "12_hexa_cell_grid_ger.shp")
```

```{r 12hexaplot}
grid_12hexacell <- st_read(dsn = "/Users/dana/Documents/R/PHD/data_maps/grids/12hexacell", 
                      layer = "12_hexa_cell_grid_ger")
grid_12hexacell <- grid_12hexacell %>%
  mutate(centroid = st_centroid(geometry))

ggplot() +
  geom_sf(data = grid_12hexacell, 
          aes(geometry = geometry),
          fill = "white") +
  geom_sf_text(data = grid_12hexacell, 
               aes(geometry = centroid, label = cell_id), 
               size = 3, 
               hjust = 0.5, # Center horizontally
               vjust = 0.5  # Center vertically
  ) +
  theme_minimal() +
  theme(axis.title.x = element_blank(), 
        axis.title.y = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        panel.grid.major = element_blank(),
        plot.title = element_text(size = 10, family = "Optima", hjust = 0.5)
  ) +
  ggtitle("12 Hexa Cell Grid")
```


## Voronoi Grid

Voronoi is meant to be an adaptive grid, so it makes cells smaller where there is more data. 

```{r voronoiprep}
# Get the token and location data to prepare the dataframe for the voronoi generation
token_at_location <- read.csv(file = './data_ling/tokens_at_location.csv')
longlat <- read.csv(file = './data_maps/germany_geo.csv')
merged_tokens <- merge(token_at_location, longlat, by.x = "city", by.y = "City")
tokens_filtered <- merged_tokens %>%
  dplyr::filter(n > 10000) %>%
  dplyr::filter(lon > 0) %>%
  na.omit()

# Assume you have a data frame `data_points` with columns `longitude` and `latitude`
# Convert your data points to an sf object for Voronoi generation
data_points_sf <- st_as_sf(tokens_filtered, coords = c("lon", "lat"), crs = 4326)

# Generate the Voronoi diagram using these points
# This is the default one
voronoi_cells <- st_voronoi(st_union(data_points_sf))

# This changes the size of the cells by adding tolerance
voronoi_cells <- st_voronoi(st_union(data_points_sf), dTolerance = 4)
voronoi_cells <- st_voronoi(st_union(data_points_sf), dTolerance = 5)
voronoi_cells <- st_voronoi(st_union(data_points_sf), dTolerance = 6)

# Convert Voronoi diagram into an sf object for further processing
voronoi_sf <- st_as_sf(st_cast(voronoi_cells))

# Clip the Voronoi cells to the boundary of Germany
germany_boundary <- st_union(ger)  # Assuming `ger` is the spatial object of Germany
voronoi_clipped <- st_intersection(voronoi_sf, germany_boundary)

# Add a unique identifier to each Voronoi cell
voronoi_clipped$cell_id <- seq_len(nrow(voronoi_clipped))
```

```{r voronoi_plot}
# Plot the Voronoi diagram overlaid with the German boundary to visualize the cells
ggplot() +
  # Plot the clipped Voronoi cells
  geom_sf(data = voronoi_clipped, fill = "white", color = "black") + 
  # Plot the original data points
  geom_sf(data = data_points_sf, color = "plum4", size = 1) + 
  theme_minimal() +
  ggtitle("Voronoi Diagram Over Germany") +
  theme(axis.title.x = element_blank(), 
        axis.title.y = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        panel.grid.major = element_blank(),
        plot.title = element_text(size = 10, family = "Optima", hjust = 0.5))
```

```{r voronoi_plot2}
# Create the plot
ggplot() +
  # Plot the clipped Voronoi cells
  geom_sf(data = voronoi_clipped, fill = "white", color = "black") +
  geom_sf_text(data = voronoi_clipped, 
               aes(label = cell_id),  # Use cell_id for labels
               size = 2,  # Smaller size to reduce clutter
               check_overlap = TRUE,  # Prevent overlapping labels
               position = position_nudge(y = 0.02)) +  # Adjust position if necessary
  theme_minimal() +
  ggtitle("Voronoi Diagram Over Germany") +
  theme(axis.title.x = element_blank(), 
        axis.title.y = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        panel.grid.major = element_blank(),
        plot.title = element_text(size = 10, family = "Optima", hjust = 0.5))
```


## Max Point Analysis for Regular Grid

```{r voronoi_check}
# Assuming max_point1 contains the coordinates of interest
max_point1 <- mean_pred_ger

# Function to check if the point is within a Voronoi cell
check_voronoi_region <- function(max_point, voronoi_grid) {
  # Create a data frame for the point's coordinates
  max_point_df <- data.frame(
    longitude = c(max_point$lon),
    latitude = c(max_point$lat)
  )
  
  # Convert to an sf object
  point_sf <- st_as_sf(max_point_df, coords = c("longitude", "latitude"), crs = st_crs(voronoi_grid))
  
  # Check if the point is within any Voronoi cell
  point_within_cell <- st_contains(voronoi_grid, point_sf, sparse = FALSE)
  
  # Find which cell the point is within
  cell_index <- which(point_within_cell, arr.ind = TRUE)
  
  # Initialize result list
  result <- list()
  
  if (length(cell_index) > 0) {
    result$cell_id <- voronoi_grid$cell_id[cell_index[1]]  # Get the cell ID
  } else {
    result$cell_id <- "Outside any cell"
  }
  
  print(result)  # Print result for inspection
  return(result)  # Return the result
}

# Check which Voronoi cell the point lies in
grid_result <- check_voronoi_region(max_point1, voronoi_clipped)
print(grid_result)
```
