---
title: "First Maps"
author: "Dana"
date: "29/03/2022"
output: html_document
toc: true
toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## First Maps

In this markdown I will summarise everything that is needed to create the GSA map, to plot data points onto it and to count instances of words per location in the corpus. I start by loading all necessary libraries and reading in the csv file for the corpus.

```{r prep}
# Load the necessary libraries for data manipulation and mapping

# 'stringr' is used for string operations, including regex functions for pattern matching and counting
library(stringr) 

# 'dplyr' provides functions for data manipulation and is commonly used for data wrangling tasks
library(dplyr) 

# 'tidyverse' is a collection of packages for data science that includes ggplot2, dplyr, tidyr, and others
# It is commonly used for data manipulation, visualization, and transformation tasks
library(tidyverse) 

# 'sp' is used for spatial data analysis and provides classes and methods for handling spatial objects
# It also supports mapping functionalities
library(sp) 

# 'rworldmap' is a package that provides functions for visualizing data on world maps, making it easy to create maps with spatial data
library(rworldmap) 

# Read in a CSV file containing training data, likely related to linguistic data, and store it in a data frame named 'corpus'
# The file path indicates that this file is located in the "data_ling" directory
corpus <- read.csv(file = "./data_ling/train_70.csv")
```

### Count Function

The next step is creating the function that counts instances of words per city. There are two versions of this function, counting two variants or just a simple count of a single feature. This is the variant counting:

```{r count_double}
# This function takes two words and produces the proportion of occurrences of the second word relative to the first.
# The first input is treated as the standard against which the second is measured.
regcount <- function(corp, word1, word2, case = FALSE, neg = FALSE) {
  
  # Prepare the first word for regex matching:
  # The '\\b' ensures that we match the whole word only (word boundaries).
  word1 <- paste('\\b', word1, '\\b', sep = "") 
  word1 <- regex(word1, ignore_case = !case) # Create a regex pattern for word1, considering case sensitivity based on the 'case' parameter.
  
  # Prepare the second word similarly.
  word2 <- paste('\\b', word2, '\\b', sep = "") 
  word2 <- regex(word2, ignore_case = !case) # Create a regex pattern for word2.

  # Extract indexes of occurrences for word1 and word2 in the text corpus:
  # 'str_which' identifies the positions in 'corp' where the pattern matches.
  hits1 <- str_which(string = corp, pat = word1, negate = neg)   
  hits2 <- str_which(string = corp, pat = word2, negate = neg)   

  # Extract the cities associated with the occurrences of word1 and word2:
  cities1 <- corpus$location[hits1] 
  cities2 <- corpus$location[hits2] 
  
  # Create frequency tables for each city:
  c1 <- as.data.frame(table(cities1)) # Frequency table for occurrences of word1.
  c2 <- as.data.frame(table(cities2)) # Frequency table for occurrences of word2.
  
  # Merge the frequency tables on city names:
  dtable <- merge(c1, c2, by.x = 'cities1', by.y = 'cities2', all = TRUE)

  # Replace any NA values in the merged table with 0:
  dtable[is.na(dtable)] <- 0
  
  # Calculate the proportion of occurrences of word1 relative to the total occurrences of both words:
  # 'Freq.x' corresponds to word1 occurrences, 'Freq.y' to word2.
  dtable$prop <- 100 * (dtable$Freq.x / (dtable$Freq.x + dtable$Freq.y))
  
  # Prepare the output data structure.
  output <- c()
  output <- dtable
  
  # Return the results containing city names, frequencies, and proportions.
  return(output)
}
```

This is the single counting:

```{r count_single}
# This function counts occurrences of a specified word in a given corpus 
# and groups the results by city.
regcount <- function(corp, word1, case = FALSE, neg = FALSE) {
  
  # Prepare the regex pattern for the specified word:
  # The '\\b' ensures that the word matches exactly (whole word).
  word1 <- paste('\\b', word1, '\\b', sep = "") 
  word1 <- regex(word1, ignore_case = !case) # Create regex pattern for word1, ignoring case based on the 'case' parameter.
  
  # Extract indexes of occurrences of word1 in the text corpus:
  # 'str_which' identifies the positions in 'corp' where the pattern matches.
  hits1 <- str_which(string = corp, pat = word1, negate = neg)   
  
  # Get the cities associated with the occurrences of word1:
  cities1 <- corpus$location[hits1] 
  
  # Create a frequency table that counts occurrences of word1 per city:
  c1 <- as.data.frame(table(cities1))
  
  # Replace any NA values in the frequency table with 0:
  c1[is.na(c1)] <- 0
  
  # Prepare the output data structure.
  output <- c()
  output <- c1
  
  # Return the results containing city names and frequencies of word1.
  return(output)
}
```

### Counting words

The next chunk first counts two words in the corpus and assigns the counts to reg_output. Reg_output then gets column names so that it can be merged with the longitudes and latitudes of the cities in the next step. 

```{r count_execution_double}
# Call the regcount function to analyze occurrences of words in the corpus.
# The 'corp' argument specifies the message column from the 'corpus', 
# The 'case' parameter is set to FALSE to ignore case sensitivity during the matching process.
reg_output <- regcount(corp = corpus$message, word1 = "zur schule", word2 = "in die schule", case = FALSE)

# Rename the columns of the resulting data frame for clarity:
# "City" will hold city names,
# "count1" will represent the count of occurrences of word 1,
# "count2" will represent the count of occurrences of word 2,
# "Proportion" will represent the proportion of occurrences for analysis.
colnames(reg_output) <- c("City", "count1", "count2", "Proportion")
```

For a single count, this is the chunk:

```{r count_execution_single}
# Call the regcount function again, this time to count occurrences of a word.
# Here, only 'word1' is specified. The function will analyze the same corpus's message column.
# The 'case' parameter remains FALSE to ensure case insensitivity.
reg_output <- regcount(corp = corpus$message, word1 = "der", case = FALSE)

# Rename the columns of the resulting data frame for clarity:
# "City" will hold city names, and "Frequency" will represent the count of occurrences of "der".
colnames(reg_output) <- c("City", "Frequency")
```


### Coordinates

In this step the coordinates for all cities are loaded and can then be merged with the counts. 

```{r coordinates}
# Read the CSV file containing geographical coordinates and city names.
# The file 'gsa_geo_filtered.csv' contains the longitudes and latitudes of various cities,
# which will be used for mapping or further geographical analysis.
longlat <- read.csv(file = './data_maps/gsa_geo_filtered.csv')

# Merge the 'reg_output' data frame, which contains city names and word counts, 
# with the 'longlat' data frame, which contains geographical coordinates.
# The merge is performed using the 'City' column from both data frames as the key.
# The result is stored in the 'merger' data frame, which now includes both word counts 
# and the corresponding geographical coordinates (latitude and longitude) for each city.
merger <- merge(reg_output, longlat, by.x = "City", by.y = "City")
```


### Initial Map

To create the outlines of the GSA based on coordinates, the following chunk gets a world map and then extracts the polygons for the GSA. Those coordinates then get stored in DACH_coord.

```{r worldmap}
# Retrieve a world map from the rworldmap package, which includes geographic information for all countries.
worldMap <- getMap()

# Define a vector containing the names of the DACH countries (Germany, Austria, Switzerland).
DACH <- c("Germany", "Austria", "Switzerland")

# Find the indices of the DACH countries in the world map data frame.
# The `which` function is used to get the indices of rows where the NAME column matches the DACH vector.
DACH_map <- which(worldMap$NAME %in% DACH)

# Create a list of data frames containing the coordinates of the polygons for each DACH country.
DACH_coord <- lapply(DACH_map, function(i) {
  # Extract the coordinates of the first polygon of the country.
  df <- data.frame(worldMap@polygons[[i]]@Polygons[[1]]@coords)
  
  # Add a new column to the data frame to specify the region (country name).
  df$region = as.character(worldMap$NAME[i])
  
  # Rename the columns for clarity: "long" for longitude, "lat" for latitude, and "region" for country name.
  colnames(df) <- list("long", "lat", "region")
  
  # Return the data frame containing the coordinates and region.
  return(df)
})

# Combine all individual data frames into a single data frame using do.call with rbind.
DACH_coord <- do.call("rbind", DACH_coord)

# OR alternative method using the rnaturalearth package
# Obtain the outlines of the DACH countries as a simple features (sf) object.
gsa_outline <- ne_countries(country = c("Austria", "Germany", "Switzerland"), returnclass = "sf", scale = "large")

# Select only the geometry column from the obtained outlines, which contains the spatial features of the countries.
gsa_plot <- gsa_outline %>% dplyr::select(geometry) # just the outlines of the three countries

# Convert the sf object to a Spatial object, which is another common format for spatial data in R.
gsa_spatial <- as_Spatial(gsa_plot) # new object as spatial instead of sf
```


### Plotting Counts

The next step then is the mapping of the coordinates with their counts using ggplot2. First, I take the GSA coordinates and create a blank map. I define which range of coordinates gets shown (coord_map) and set the theme to minimal. I then tell ggplot to use the merger data and plot points on my blank map using coordinates. The colour of the points is based on the proportion of use of the first (or main) word in the count. City labels can be added. scale_color_gradient colours the gradient of the proportion (and the points). Different color filling is needed for categorical data. Labs sets the title of the legend. ggtitle sets the title of the whole plot. The last theme element gets rid of grid lines and marks on the axis, including axis title.
This is the map for variant / proportion mapping.

```{r mapping_double}
# Begin creating a ggplot object
ggplot() + 

  # Add polygons for the DACH countries using the DACH_coord data frame
  geom_polygon(data = DACH_coord, 
               aes(x = long, y = lat, group = region),  # Map longitudes and latitudes with group defined by region
               colour = "black",       # Set the border color of the polygons to black
               size = 0.1,            # Set the border size for better visibility
               fill = "snow3") +      # Fill the polygons with a light snow color

  # Set the coordinate system for the map with specific limits on x and y axes
  coord_map(xlim = c(4.5, 17),      # Set longitude limits for the plot
            ylim = c(45.5, 55)) +   # Set latitude limits for the plot

  # Use a minimal theme for cleaner visuals
  theme_minimal() +  

  # Add points to the map based on data from the merger data frame
  geom_point(data = merger, 
             aes(x = lon, y = lat,                # Map longitude and latitude for points
                 col = Proportion,                 # Color points based on the 'Proportion' variable
                 size = (count1 + count2)),      # Size of points based on the sum of count1 and count2
             alpha = 0.9) +                      # Set the transparency of the points

  # Uncomment the line below to display city labels on the map
  # geom_text(data = merger, aes(x = lng, y = lat, label = city), hjust = 0.1, vjust = 0.1, size = 2) +

  # Suppress the size legend in the plot
  guides(size = FALSE) +

  # Create a color gradient for the points from low to high proportions
  scale_color_gradient(low = "seagreen3", high = "mediumpurple3") +

  # Set labels for the size legend
  labs(size = "Anteil") +

  # Add a title to the plot
  ggtitle("nächstes/n monat") +

  # Customize theme elements to improve plot appearance
  theme(axis.title.x = element_blank(),   # Remove x-axis title
        axis.title.y = element_blank(),   # Remove y-axis title
        axis.text.x = element_blank(),     # Remove x-axis text labels
        axis.text.y = element_blank(),     # Remove y-axis text labels
        panel.grid.major = element_blank(), # Remove major grid lines
        plot.title = element_text(hjust = 0.5)) # Center the plot title
```

```{r mapping_double_darkmode}
# Begin creating a ggplot object
ggplot() + 
  
  # Add spatial features (geometries) for the DACH countries using the gsa_plot data frame
  geom_sf(data = gsa_plot, 
          aes(geometry = geometry),  # Map geometries directly from the sf object
          color = "snow2",           # Set the border color of the shapes to a light snow color
          fill = "#303030",          # Fill the shapes with a dark gray color
          size = 0.5) +              # Set the size of the borders for visibility
  
  # Add points to the map based on data from the merger data frame
  geom_point(data = merger, 
             aes(x = lon, y = lat,               # Map longitude and latitude for points
                 col = Proportion,                # Color points based on the 'Proportion' variable
                 size = (Frequency.x + Frequency.y)),  # Size of points based on the sum of frequency counts
             alpha = 0.9) +                     # Set the transparency of the points
  
  # Add text labels for cities on the map using cities_sf data
  geom_sf_text(data = cities_sf, 
               aes(label = City),               # Use the 'City' variable for labeling
               size = 2.5,                      # Set the size of the text labels
               nudge_x = 0,                     # No horizontal nudging
               nudge_y = -0.15,                 # Slightly nudge labels downward for better visibility
               family = "Optima",               # Specify font family for the labels
               color = "snow2") +               # Set label color to light snow color
  
  # Add markers (crosses) for cities using the same cities_sf data
  geom_sf(data = cities_sf, 
          aes(geometry = geometry),         # Map the geometries for city markers
          shape = 4,                       # Set shape to a cross (type 4)
          color = "snow2") +               # Set color of the markers to light snow
  
  # Use a minimal theme for cleaner visuals
  theme_minimal() +  
  
  # Suppress the size legend in the plot
  guides(size = FALSE) +
  
  # Create a color gradient for the points from low to high proportions
  scale_color_gradient(low = "blue", high = "firebrick1") +
  
  # Set a label for the color legend
  labs(col = "schau / guck") +  # add legend heading
  
  # Customize various theme elements for better visual appeal
  theme(axis.title.x = element_blank(),          # Remove x-axis title
        axis.title.y = element_blank(),          # Remove y-axis title
        axis.text.x = element_blank(),            # Remove x-axis text labels
        axis.text.y = element_blank(),            # Remove y-axis text labels
        legend.position = c(0.90, 0.65),          # Position the legend within the plot area
        legend.title = element_text(size = 6, color = "snow", family = "Optima"),  # Customize legend title font
        legend.text = element_text(size = 6, color = "snow", family = "Optima", hjust = 1),  # Customize legend text font
        legend.title.align = 0,                    # Align legend title to the left
        panel.background = element_rect(fill = "#303030", color = "#303030"),  # Set panel background to dark gray
        plot.background = element_rect(fill = "#303030", color = "#303030"),   # Set overall plot background to dark gray
        legend.key.size = unit(0.3, "cm"),        # Set the size of legend keys
        legend.key.width = unit(0.4, "cm"),       # Set the width of legend keys
        panel.grid.major = element_blank())        # Remove major grid lines
```


```{r mapping_double_lightmode_updated}
# Begin creating a ggplot object
ggplot() + 

  # Add spatial features (geometries) for the DACH countries using the gsa_plot data frame
  geom_sf(data = gsa_plot, 
          aes(geometry = geometry),  # Map geometries directly from the sf object
          color = "black",           # Set the border color of the shapes to black
          fill = "snow1",           # Fill the shapes with a light snow color
          size = 0.5) +             # Set the size of the borders for visibility

  # Add points to the map based on data from the merger data frame
  geom_point(data = merger, 
             aes(x = lon, y = lat,                # Map longitude and latitude for points
                 col = Proportion,                 # Color points based on the 'Proportion' variable
                 size = (Frequency.x + Frequency.y)),  # Size of points based on the sum of frequency counts
             alpha = 0.9) +                      # Set the transparency of the points

  # Use a minimal theme for cleaner visuals
  theme_minimal() +  

  # Suppress the size legend in the plot
  guides(size = FALSE) +

  # Create a color gradient for the points from low to high proportions
  scale_color_gradient(low = "seagreen3", high = "mediumpurple3") +

  # Set a label for the color legend
  labs(col = "schau / guck") +  # 'schau / guck' translates to 'look / see' in English

  # Customize various theme elements for better visual appeal
  theme(axis.title.x = element_blank(),          # Remove x-axis title
        axis.title.y = element_blank(),          # Remove y-axis title
        axis.text.x = element_blank(),            # Remove x-axis text labels
        axis.text.y = element_blank(),            # Remove y-axis text labels
        legend.position = c(0.90, 0.65),          # Position the legend within the plot area
        legend.title = element_text(size = 6, color = "black", family = "Optima"),  # Customize legend title font
        legend.text = element_text(size = 6, color = "black", family = "Optima", hjust = 1),  # Customize legend text font
        legend.title.align = 0,                    # Align legend title to the left
        legend.key.size = unit(0.3, "cm"),        # Set the size of legend keys
        legend.key.width = unit(0.4, "cm"),       # Set the width of legend keys
        panel.grid.major = element_blank())        # Remove major grid lines
```


This is the map for a single count in dark mode, followed by the light mode version.

```{r mapping_single_darkmode}
# Begin creating a ggplot object
ggplot() + 
  
  # Add spatial features (geometries) for the DACH countries using the gsa_plot data frame
  geom_sf(data = gsa_plot, 
          aes(geometry = geometry),  # Map geometries directly from the sf object
          color = "snow2",           # Set the border color of the shapes to a light snow color
          fill = "#303030",          # Fill the shapes with a dark gray color
          size = 0.5) +              # Set the size of the borders for visibility
  
  # Add points to the map based on data from the merger data frame
  geom_point(data = merger, 
             aes(x = lon, y = lat,          # Map longitude and latitude for points
                 size = Frequency),           # Size of points based on the 'Frequency' variable
             col = "purple",                 # Set the color of the points to purple
             alpha = 0.9) +                  # Set the transparency of the points
  
  # Add text labels for cities on the map using cities_sf data
  geom_sf_text(data = cities_sf, 
               aes(label = City),             # Use the 'City' variable for labeling
               size = 2.5,                    # Set the size of the text labels
               nudge_x = 0,                   # No horizontal nudging
               nudge_y = -0.15,               # Slightly nudge labels downward for better visibility
               family = "Optima",             # Specify font family for the labels
               color = "snow2") +             # Set label color to light snow color
  
  # Add markers (crosses) for cities using the same cities_sf data
  geom_sf(data = cities_sf, 
          aes(geometry = geometry),       # Map the geometries for city markers
          shape = 4,                     # Set shape to a cross (type 4)
          color = "snow2") +             # Set color of the markers to light snow
  
  # Use a minimal theme for cleaner visuals
  theme_minimal() +  
  
  # Suppress the size legend in the plot
  guides(size = "none") +            # 'none' disables the legend for size
  
  # Customize various theme elements for better visual appeal
  theme(axis.title.x = element_blank(),          # Remove x-axis title
        axis.title.y = element_blank(),          # Remove y-axis title
        axis.text.x = element_blank(),            # Remove x-axis text labels
        axis.text.y = element_blank(),            # Remove y-axis text labels
        panel.background = element_rect(fill = "#303030", color = "#303030"),  
        # Set panel background to dark gray
        plot.background = element_rect(fill = "#303030", color = "#303030"),   
        # Set overall plot background to dark gray
        panel.grid.major = element_blank())        # Remove major grid lines
```

```{r mapping_single}
# Start creating a ggplot object
ggplot() + 
  
  # Add spatial features (geometries) for the DACH countries using the gsa_plot data frame
  geom_sf(data = gsa_plot, 
          aes(geometry = geometry),        # Map geometries directly from the sf object
          colour = "black",                # Set the border color of the shapes to black
          size = 0.5,                      # Set the size of the borders for visibility
          fill = "gray95") +               # Fill the shapes with a light gray color
  
  # Add points to the map based on data from the merger data frame
  geom_point(data = merger, 
             aes(x = lon, y = lat,        # Map longitude and latitude for points
                 size = Frequency),         # Size of points based on the 'Frequency' variable
             colour = "darkorchid4",       # Set the color of the points to a dark orchid color
             alpha = 0.7) +                # Set the transparency of the points to 70%
  
  # Add text labels for cities on the map using cities_sf data
  geom_sf_text(data = cities_sf, 
               aes(label = City),           # Use the 'City' variable for labeling
               size = 2.5,                  # Set the size of the text labels
               nudge_x = 0,                 # No horizontal nudging
               nudge_y = -0.15,             # Slightly nudge labels downward for better visibility
               family = "Optima",           # Specify font family for the labels
               color = "black") +           # Set label color to black
  
  # Add markers (crosses) for cities using the same cities_sf data
  geom_sf(data = cities_sf, 
          aes(geometry = geometry),       # Map the geometries for city markers
          shape = 4,                     # Set shape to a cross (type 4)
          color = "black") +             # Set color of the markers to black
  
  # Use a minimal theme for cleaner visuals
  theme_minimal() +  
  
  # Suppress the size legend in the plot
  guides(size = "none") +            # 'none' disables the legend for size
  
  # Customize various theme elements for better visual appeal
  theme(axis.title.x = element_blank(),          # Remove x-axis title
        axis.title.y = element_blank(),          # Remove y-axis title
        axis.text.x = element_blank(),            # Remove x-axis text labels
        axis.text.y = element_blank(),            # Remove y-axis text labels
        panel.grid.major = element_blank())        # Remove major grid lines
```


Maps can then be easily exported using the following code chunk.

```{r export}
# Save the current ggplot object to a PNG file
ggsave("./output/full_german1.png",      # Specify the file path and name for the saved plot
       width = 6.5,                       # Set the width of the saved image in inches
       height = 5.5)                      # Set the height of the saved image in inches
```

