---
title: "Shape Files Mapping"
author: "Dana"
date: '2022-10-24'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Mapping with Shape Files

This RMD is about using the shape files for the German-speaking area and filling the shapes given a frequency. 
I start by loading the libraries. 

```{r libs}
library(sf)
library(tidyverse)
library(sp)
library(rworldmap)
library(dplyr)
library(RColorBrewer)
```

Then I load in the shape files. To know which shape files work is a bit of trial and error. I have used shape files that correspond to postal codes for Switzerland and Germany, whereas in Austria the basis for merging is Gemeindekennziffern, which roughly translates to administrative district number. 

```{r load shapes}
germany <- st_read(dsn="/rds/projects/g/grievej-german-dialect-profiling/shape/Germany/plz-3stellig/", layer="plz-3stellig")
austria <- st_read(dsn="/rds/projects/g/grievej-german-dialect-profiling/shape/Austria/Gemeinden/", layer="Gemeinden_250")
switzerland <- st_read(dsn="/rds/projects/g/grievej-german-dialect-profiling/shape/Switzerland/PC/", layer="PLZO_PLZ")
EU <- st_read(dsn="/rds/projects/g/grievej-german-dialect-profiling/shape/EU/Try/", layer="NUTS_RG_20M_2021_3035")
#plot(germany)
#plot(austria)
#plot(switzerland)
```

I change the projection to the same one for all of the shape files. This ensures that they can all be mapped together, as different projections would project the shapes differently. WGS84 is one of the standard projections, which refers to "a datum featuring coordinates that change with time. WGS84 is defined and maintained by the United States National Geospatial-Intelligence Agency (NGA)."

```{r projection}
germany <- st_transform(germany, CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"))
austria <- st_transform(austria, CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"))
switzerland <- st_transform(switzerland, CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"))
EU <- st_transform(EU, CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"))
```

I then get the outlines I want to plot around the filled shapes. I do this by selecting the relevant country names / abbreviations and filtering the EU shape files for these countries. 

```{r gsa outline}
gsa_string <- c("DE", "AT", "CH")
gsa <- filter(EU, CNTR_CODE %in% gsa_string)
gsa_outline <- filter(gsa, LEVL_CODE == 0)
#plot(gsa_outline)
```

Now I load in the name & post code match files. 

```{r load post code}
ger_cities <- read.csv(file = '/rds/projects/g/grievej-german-dialect-profiling/shape/germany.csv', colClasses='character')
aus_cities <- read.csv(file = '/rds/projects/g/grievej-german-dialect-profiling/shape/austria.csv', colClasses='character')
s_cities <- read.csv(file = '/rds/projects/g/grievej-german-dialect-profiling/shape/switzerland.csv', colClasses='character')
```

Now I can merge the shape files with the city name information. 

```{r merge}
merger_ger <- as.data.frame(merge(ger_cities, germany, by.x ="PLZ3_Ger", by.y = "plz", all.x = TRUE))
merger_aus <- merge(aus_cities, austria, by.x ="Gemeinde_Aus", by.y = "GKZ", all.x = TRUE)
merger_s <- merge(s_cities, switzerland, by.x ="PLZ4_Swi", by.y = "PLZ", all.x = TRUE)

```

With this, I can make simple plots. I could have made a plot without having the names in the shape files and just use the shape files, but then I couldn't combine it with the Jodel data.

```{r simple plots}
#Austria plots quickly, just the areas where there is data
ggplot(data = merger_aus, aes(geometry = geometry)) +
  geom_sf()

#Switzerland plots quickly too, just the areas where there is data
ggplot(data = merger_s, aes(geometry = geometry)) +
  geom_sf()

#Germany takes quite long
ggplot(data = merger_ger, aes(geometry = geometry)) +
  geom_sf()

#first combined plot
ggplot() +
  geom_sf(data = gsa_outline, aes(geometry = geometry), colour="black", fill=NA) + 
  geom_sf(data = merger_ger, aes(geometry = geometry)) +
  geom_sf(data = merger_aus, aes(geometry = geometry)) +
  geom_sf(data = merger_s, aes(geometry = geometry)) 
```

Then I make the map more minimal.

```{r minimal}
#more minimal combined plot
ggplot() +
  geom_sf(data = gsa_outline, aes(geometry = geometry), colour="black", fill=NA) +
  theme_minimal() +
  theme(axis.title.x = element_blank(), 
        axis.title.y = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        panel.grid.major = element_blank(),
        plot.title = element_text(hjust = 0.5)) +
  geom_sf(data = merger_ger, aes(geometry = geometry), colour="black", fill=NA) +
  geom_sf(data = merger_aus, aes(geometry = geometry), colour="black", fill=NA) +
  geom_sf(data = merger_s, aes(geometry = geometry), colour="black", fill=NA)
```

To plot something I counted, I get the counts from the matrix data. 

```{r count function}
#get word of interest and merge with geo data
corpus <- read.csv(file = './data_ling/full_matrix_for_filtering.csv') 
one_word <- corpus %>% filter(word == "ich")
colnames(one_word) <- c("Token", "City", "Frequency")
```

And now merging the counts and the shape files that already contained the city names.

```{r merge count}
plot_ger <- merge(merger_ger, one_word, by ="City", all.x = TRUE)
plot_aus <- merge(merger_aus, one_word, by ="City", all.x = TRUE)
plot_swiz <- merge(merger_s, one_word, by ="City", all.x = TRUE)
```

Finally, I can produce the map.

```{r final plot}
#dev.off()
ggplot() +
  geom_sf(data = gsa_outline, aes(geometry = geometry), colour="black", fill=NA, size = 0.1) +
  theme_minimal() +
  theme(axis.title.x = element_blank(), 
        axis.title.y = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        panel.grid.major = element_blank(),
        plot.title = element_text(hjust = 0.5)) +
  geom_sf(data = plot_aus[!is.na(plot_aus$Frequency),], 
          aes(geometry = geometry, fill=Frequency), size = 0.05) + 
  geom_sf(data = plot_swiz[!is.na(plot_swiz$Frequency),], 
          aes(geometry = geometry, fill=Frequency), size = 0.05) +
  geom_sf(data = plot_ger[!is.na(plot_ger$Frequency),], 
          aes(geometry = geometry, fill=Frequency), size = 0.05) +
  scale_fill_gradient2(midpoint = 0) +
  ggtitle("Frequency of 'Matura' in the GSA") +
  labs(fill='Matura') 
```





