---
title: "Relative Frequency Mapping"
author: "Dana"
date: '2022-11-23'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Mapping with Relative Frequency

This Rmd is an updated version of the mappings in 3_first_maps to include the relative frequency, but also to no longer rely on the reg_count() function, but base all counts on the matrix. 

```{r libs}
# libraries
library(stringr) # regex
library(dplyr) # regex
library(tidyverse) # mapping
library(rworldmap) # mapping

# data standard
corpus <- read.csv(file = './data_ling/full_matrix_for_filtering.csv') 
token_at_location <- read.csv(file = './data_ling/tokens_at_location.csv')
colnames(token_at_location) <- c("City", "Tokencount") 
longlat <- read.csv(file = './data_maps/gsa_geo_filtered.csv')

# maybe
corpus_full <- read.csv(file = './data_ling/train_70.csv')
matrix <- read.csv(file = './data_ling/gsa_geo_filtered.csv/10k_matrix.csv', header = TRUE) 
matrix_geo <- read.csv(file = './data_ling/gsa_geo_filtered.csv/10k_matrix_geo.csv', header = TRUE) 
```

Instead of counting, I can just filter the matrix for the word I am interested in. 

```{r filtering}
#get word of interest and merge with geo data
one_word <- corpus %>% filter(word == "ich")
colnames(one_word) <- c("Token", "City", "Frequency")
merger_one <- merge(one_word, longlat, by.x ="City", by.y = "City")
colnames(merger_one) <- c("City", "Token", "Frequency", "lon", "lat")
```

Now I can add the relative frequency. For that I merge with the counts of all tokens at all locations. This information is merged with the data and the coordinates details. With that in one object, the relative frequency can be calculated. 

```{r relative frequency}
merger <- merge(merger_one, token_at_location, by.x ="City", by.y = "City")
merger$relfreq <- (merger$Frequency/merger$Tokencount)
merger$relfreq1000 <- (merger$relfreq*1000)
colnames(merger) <- c("City", "Token", "Frequency", "lon", "lat", "TokenCount", "RelativeFrequency", "RelativeFrequencyThousand")
```

Then the usual map preparation is done:

```{r map prep}
worldMap <- getMap()
DACH <- c("Germany", "Austria", "Switzerland")
DACH_map <- which(worldMap$NAME%in%DACH)
DACH_coord <- lapply(DACH_map, function(i){
  df <- data.frame(worldMap@polygons[[i]]@Polygons[[1]]@coords)
  df$region =as.character(worldMap$NAME[i])
  colnames(df) <- list("long", "lat", "region")
  return(df)
})
DACH_coord <- do.call("rbind", DACH_coord)
```

And now I can map the relative frequency:

```{r map plot}
ggplot() + geom_polygon(data = DACH_coord, 
                        aes(x = long, y = lat, group = region),
                        colour = "black", 
                        size = 0.1, 
                        fill = "gray95") + 
  coord_map(xlim = c(4.5, 17),  
            ylim = c(45.5, 55)) + 
  theme_minimal() +  
  geom_point(data = merger, 
             aes(x = lon, y = lat, size = RelativeFrequencyThousand),
             colour = "darkorchid4",
             alpha = 0.7)  +
  #geom_text(data = merger, aes(x = lng, y = lat, label=city), hjust=0.1, vjust=0.1, size = 2) +
  guides(color = "none") +
  labs(size="Relative\nFrequency") +
  ggtitle("'nix' in the GSA") +
  theme(axis.title.x = element_blank(), 
        axis.title.y = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        panel.grid.major = element_blank(),
        plot.title = element_text(hjust = 0.5))
```

The plot can be exported using ggsave.

```{r export}
ggsave("./output/mistkübel.png", width = 6.5, height = 5.5)
```

I can also plot just the frequencies: 

```{r raw frequency}
ggplot() + geom_polygon(data = DACH_coord, 
                        aes(x = long, y = lat, group = region),
                        colour = "black", 
                        size = 0.1, 
                        fill = "gray95") + 
  coord_map(xlim = c(4.5, 17),  
            ylim = c(45.5, 55)) + 
  theme_minimal() +  
  geom_point(data = merger_one, 
             aes(x = lon, y = lat, size = Frequency),
             colour = "darkorchid4",
             alpha = 0.7)  +
  #geom_text(data = merger, aes(x = lng, y = lat, label=city), hjust=0.1, vjust=0.1, size = 2) +
  guides(color = "none") +
  labs(size="Frequency") +
  ggtitle("'ich' in the GSA") +
  theme(axis.title.x = element_blank(), 
        axis.title.y = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        panel.grid.major = element_blank(),
        plot.title = element_text(hjust = 0.5))
```


