---
title: "Updated Maps"
author: "Dana Roemling"
date: '2023-03-14'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r libs}
library(stringr) 
library(dplyr) 
library(tidyverse) 
library(rnaturalearth)  
```

## New GSA outline

There is a new and easier way to create the GSA outline, which is based on the ne_countries() function. 

```{r outline}
gsa_outline <- ne_countries(country = c("Austria", "Germany", "Switzerland"), returnclass="sf", scale = "large")
```

## Preparation

This chunk needs to run for the altered mapping to work (just the usual prep.).

```{r prep}
# data standard
corpus <- read.csv(file = './data_ling/full_matrix_for_filtering.csv') 
token_at_location <- read.csv(file = './data_ling/tokens_at_location.csv')
colnames(token_at_location) <- c("City", "Tokencount") 
longlat <- read.csv(file = './data_maps/gsa_geo_filtered.csv')
longlat_small <- read.csv(file = './data_maps/gsa_geo_filtered_nosmall22.csv')
longlat_small <- read.csv(file = './data_maps/gsa_geo_filtered_nosmall85.csv')

# get word of interest
one_word <- corpus %>% filter(word == "mistkübel")
colnames(one_word) <- c("Token", "City", "Frequency")

# merge word with geo data
merger_one <- merge(one_word, longlat, by.x ="City", by.y = "City")
colnames(merger_one) <- c("City", "Token", "Frequency", "lon", "lat")

# add relative frequency
merger <- merge(merger_one, token_at_location, by.x ="City", by.y = "City")
merger$relfreq <- (merger$Frequency/merger$Tokencount)
merger$relfreq1000 <- (merger$relfreq*1000)
colnames(merger) <- c("City", "Token", "Frequency", "lon", "lat", "TokenCount", "RelativeFrequency", "RelativeFrequencyThousand")

```

## Plotting

Instead of using geom_polygon() I now need to use geom_sf() to create the map. But the rest of the mapping, like the point layer, remains the same.

```{r plot}
ggplot(data = gsa_outline,
       colour = "black", 
       size = 0.1, 
       fill = "gray95") + 
  geom_sf() + 
  theme_minimal() +  
  geom_point(data = merger, 
             aes(x = lon, y = lat, size = RelativeFrequencyThousand),
             colour = "darkorchid4",
             alpha = 0.7)  +
  #geom_text(data = merger, aes(x = lng, y = lat, label=city), hjust=0.1, vjust=0.1, size = 2) +
  guides(color = "none") +
  labs(size="Relative\nFrequency") +
  ggtitle("'möchten' in the GSA") +
  theme(axis.title.x = element_blank(), 
        axis.title.y = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        panel.grid.major = element_blank(),
        plot.title = element_text(hjust = 0.5))
```

# Plotting Comparison

I can also do the alternation plots with the new GSA outline. A quick prep to get the values first.

```{r prep 2}
one_word <- corpus %>% filter(word == "mülleimer")
colnames(one_word) <- c("Token", "City", "Frequency")
second_word <- corpus %>% filter(word == "mistkübel")
colnames(second_word) <- c("Token", "City", "Frequency")

# combine into one and calculate proportion
merged <- merge(one_word, second_word, by ="City", all=T)
# the next two lines make sure that ALL instances get plotted.
merged$Frequency.x[is.na(merged$Frequency.x)] <- 0
merged$Frequency.y[is.na(merged$Frequency.y)] <- 0
merged$Proportion <- 100* (merged$Frequency.x / (merged$Frequency.x + merged$Frequency.y))

# merge with geo data
merger <- merge(merged, longlat, by.x ="City", by.y = "City")
```


```{r plot 2}
newtitle <- expression(paste(italic("matura "), "& ", italic("abitur "), "in the GSA"))

ggplot(data = gsa_outline,
       colour = "black", 
       size = 0.1, 
       fill = "gray95") + 
  geom_sf() + 
  theme_minimal() +  
  geom_point(data = merger, 
             aes(x = lon, y = lat, col = Proportion, size = (Frequency.x+Frequency.y)), 
             alpha = 0.9)  +
  #comment out line below to lose city labels
  #geom_text(data = merger, aes(x = lng, y = lat, label=city), hjust=0.1, vjust=0.1, size = 2) +
  #scale_color_viridis(option = "magma") +
  guides(size = "none") +
  scale_color_gradient(low = "seagreen3", high = "mediumpurple3") +
  labs(size="Proportion") +
  ggtitle(newtitle) +
  theme(axis.title.x = element_blank(), 
        axis.title.y = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        panel.grid.major = element_blank(),
        legend.text = element_text(size = 6),
        legend.title = element_text(size = 9),
        plot.title = element_text(hjust = 0.4, size = 13, vjust = -2))
```
