---
title: "Updated Maps"
author: "Dana Roemling"
date: '2023-03-14'
output: html_document
toc: true
toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r libs}
# Load the stringr package for string manipulation and operations
library(stringr) 

# Load the dplyr package, which provides tools for data manipulation (e.g., filtering, selecting, mutating)
library(dplyr) 

# Load the tidyverse, a collection of packages for data science, including ggplot2, dplyr, tidyr, and others
library(tidyverse) 

# Load rnaturalearth, which provides world maps and country borders data, useful for plotting geographic data
library(rnaturalearth) 
```

## New GSA outline

There is a new and easier way to create the GSA outline, which is based on the ne_countries() function. 

```{r outline}
# Use the ne_countries function from the rnaturalearth package to retrieve country boundaries
# Specify the countries of interest: Austria, Germany, and Switzerland
# The 'returnclass="sf"' argument ensures the data is returned as an "sf" (simple features) object, 
# which is useful for spatial data manipulation and plotting in ggplot2
# The 'scale="large"' option ensures high-resolution boundary data for more detailed mapping
gsa_outline <- ne_countries(country = c("Austria", "Germany", "Switzerland"), 
                            returnclass = "sf", 
                            scale = "large")
```

## Preparation

This chunk needs to run for the altered mapping to work (just the usual prep.).

```{r prep}
# Load data from CSV files into R

# Load the main corpus data (full matrix for filtering) which contains word usage information
corpus <- read.csv(file = './data_ling/full_matrix_for_filtering.csv') 

# Load the token count at different city locations
token_at_location <- read.csv(file = './data_ling/tokens_at_location.csv')

# Rename columns in token_at_location for clarity (City and Tokencount)
colnames(token_at_location) <- c("City", "Tokencount") 

# Load geographic coordinates (longitude, latitude) for city locations from multiple CSV files
longlat <- read.csv(file = './data_maps/gsa_geo_filtered.csv')
longlat_small1 <- read.csv(file = './data_maps/gsa_geo_filtered_nosmall22.csv')
longlat_small2 <- read.csv(file = './data_maps/gsa_geo_filtered_nosmall85.csv')
longlat_small3 <- read.csv(file = './data_maps/gsa_geo_filtered_nosmall14378.csv')

# Filter the corpus for the word of interest
one_word <- corpus %>% filter(word == "adios")

# Rename columns in one_word for clarity (Token, City, Frequency)
colnames(one_word) <- c("Token", "City", "Frequency")

# Merge the filtered word data (one_word) with geographic data (longlat_small), matching by "City"
merger_one <- merge(one_word, longlat_small, by.x = "City", by.y = "City")

# Rename columns in the resulting merged data for clarity
colnames(merger_one) <- c("City", "Token", "Frequency", "lon", "lat")

# Merge the result (merger_one) with token count data (token_at_location) by "City"
merger <- merge(merger_one, token_at_location, by.x = "City", by.y = "City")

# Alternatively, perform another type of merge to include zeros for missing data (all cities) and merge again

# Merge the geographic data (longlat_small) with word data (one_word), including unmatched rows (all=TRUE)
new_join <- merge(longlat_small, one_word, by.x = "City", by.y = "City", all = TRUE)

# Select only relevant columns (City, lon, lat, Frequency) from the merged data
new_join <- new_join %>% dplyr::select("City", "lon", "lat", "Frequency")

# Remove rows with missing longitude values
new_join <- new_join %>% drop_na(lon)

# Replace all remaining NAs (missing values) with 0 in the dataset
new_join[is.na(new_join)] <- 0

# Merge the updated dataset (new_join) with token count data by "City"
merger <- merge(new_join, token_at_location, by.x = "City", by.y = "City", all.x = TRUE)

# Calculate the relative frequency of the word (adios) in each city
merger$relfreq <- (merger$Frequency / merger$Tokencount)

# Multiply relative frequency by 1000 to scale it for better interpretation
merger$relfreq1000 <- (merger$relfreq * 1000)

# Rename columns for clarity
colnames(merger) <- c("City", "Token", "Frequency", "lon", "lat", "TokenCount", "RelativeFrequency", "RelativeFrequencyThousand")

# Optionally: Drop rows where RelativeFrequency is NA
# merger <- merger %>% drop_na(RelativeFrequency)

# Optionally: Standardize (scale) the RelativeFrequency values
# merger$Scaled <- scale(merger$RelativeFrequency)

# Filter out cities with low token counts by only keeping cities with a TokenCount higher than the 1st quartile
quant <- summary(merger$TokenCount)[2]
merger <- merger %>% filter(TokenCount > quant)
```

## Plotting

Instead of using geom_polygon() I now need to use geom_sf() to create the map. But the rest of the mapping, like the point layer, remains the same.

```{r plot_1}
ggplot(data = gsa_outline,
       colour = "black",  # Set the outline color of the countries to black
       size = 0.1,        # Define the thickness of the country borders
       fill = "gray98") + # Fill the countries with a light gray color
  geom_sf() +             # Plot the geographical data as simple features (sf object)
  theme_minimal() +       # Use a minimal theme for a clean, simple background
  geom_point(data = merger, 
             aes(x = lon, 
                 y = lat, 
                 size = RelativeFrequencyThousand), # Plot points at city locations using 
             # longitude (lon) and latitude (lat) with point size representing 'RelativeFrequencyThousand'
             colour = "darkorchid4",  # Set the color of the points to dark purple
             alpha = 0.7)  +          # Set transparency of points to 70% for better visual clarity
  #geom_text(data = merger, aes(x = lng, y = lat, label=city), hjust=0.1, vjust=0.1, size = 2) + 
  # Optional: Uncomment this line above to add city labels near the plotted points
  guides(color = "none") +  # Remove the color legend from the plot (no need since points are all the same color)
  labs(size="schmäh\nRelative\nFrequency") +  # Label the point size legend with a custom title for clarity
  #ggtitle("'möchten' in the GSA") +  # Optional: Add a title to the plot (commented out)
  theme(axis.title.x = element_blank(),      # Remove the x-axis title
        axis.title.y = element_blank(),      # Remove the y-axis title
        axis.text.x = element_blank(),       # Remove x-axis text labels
        axis.text.y = element_blank(),       # Remove y-axis text labels
        panel.grid.major = element_blank(),  # Remove the major grid lines for a cleaner look
        plot.title = element_text(hjust = 0.5))  # Center the plot title (if added)
```

```{r plot_exported}
ggplot(data = gsa_outline,
       colour = "black",  # Set the country borders color to black
       size = 0.1,        # Border thickness for the country outlines
       fill = "gray98") + # Light gray fill for the countries
  geom_sf() +             # Plot the geographical shapes as simple features (sf object)
  theme_minimal() +       # Use a clean, minimalistic background for the plot
  geom_point(data = merger, 
             aes(x = lon, 
                 y = lat, 
                 size = RelativeFrequencyThousand), # Plot points at city locations using 
             # longitude (lon) and latitude (lat), with point size based on 'RelativeFrequencyThousand'
             colour = "darkorchid4",  # Points are colored dark purple
             alpha = 0.7)  +          # Set transparency of points for clarity (70% opaque)
  ggtitle(label = "'hund'",           # Add a plot title: "'hund'"
          subtitle = "median exclusion") +  # Add a subtitle: "median exclusion"
  #geom_text(data = merger, aes(x = lng, y = lat, label=city), hjust=0.1, vjust=0.1, size = 2) + 
  # Optional: Uncomment to add city labels near the points
  guides(color = "none") +            # Remove color legend as color is constant
  labs(size="Relative\nFrequency") +  # Label the size legend with "Relative Frequency"
  #ggtitle("'möchten' in the GSA") +  # Optional: Commented out alternative title
  theme(axis.title.x = element_blank(),       # Remove the x-axis title
        axis.title.y = element_blank(),       # Remove the y-axis title
        axis.text.x = element_blank(),        # Remove the x-axis text labels
        axis.text.y = element_blank(),        # Remove the y-axis text labels
        panel.grid.major = element_blank(),   # Remove the major grid lines for a cleaner plot
        legend.title = element_text(size=7, family = "Optima"),  # Set font style and size for the legend title
        legend.text = element_text(size = 5, family = "Optima"), # Set font style and size for the legend text
        plot.title = element_text(family = "Optima"),            # Set font style for the plot title
        plot.subtitle = element_text(family = "Optima"))         # Set font style for the plot subtitle
```


# Plotting Comparison

I can also do the alternation plots with the new GSA outline. A quick prep to get the values first.

```{r prep_2}
# Extract data for a word from the corpus
one_word <- corpus %>% filter(word == "werbung")
colnames(one_word) <- c("Token", "City", "Frequency")  # Rename columns to 'Token', 'City', 'Frequency'

# Extract data for a second word from the corpus
second_word <- corpus %>% filter(word == "reklame")
colnames(second_word) <- c("Token", "City", "Frequency")  # Rename columns for consistency

# Combine the two datasets based on the 'City' column
# This allows comparison of the usage of "the two words across cities
merged <- merge(one_word, second_word, by ="City", all=T)

# The next two lines ensure that missing values (NA) for either word frequency
# are replaced with zeros, ensuring all instances are accounted for in the proportion calculation
merged$Frequency.x[is.na(merged$Frequency.x)] <- 0  # Replace NA frequency with 0
merged$Frequency.y[is.na(merged$Frequency.y)] <- 0  # Replace NA frequency with 0

# Calculate the proportion of first word usage compared to the total of both words
# This is done by dividing the frequency of word one by the sum of both word frequencies
merged$Proportion <- 100 * (merged$Frequency.x / (merged$Frequency.x + merged$Frequency.y))  # Convert to percentage

# Merge the frequency data with geographical coordinates
# This adds longitude and latitude information to the merged dataset for plotting
merger <- merge(merged, longlat, by.x ="City", by.y = "City")
```


```{r plot 2}
# Define the title of the plot using expression to italicize the words and include "&"
newtitle <- expression(paste(italic("matura "), "& ", italic("abitur "), "in the GSA"))

# Create the base plot using ggplot and add data from 'gsa_outline'
ggplot(data = gsa_outline,
       colour = "black",    # Set outline color to black
       size = 0.1,          # Set size of the outline
       fill = "gray95") +   # Set fill color for the background of the map
  geom_sf() +               # Add the geographical shapes (countries, etc.)
  theme_minimal() +         # Use a minimal theme for the plot layout (clean and simple)
  
  # Add points from the 'merger' data, mapping longitude (lon) and latitude (lat)
  geom_point(data = merger, 
             aes(x = lon, 
                 y = lat, 
                 col = Proportion, 
                 size = (Frequency.x + Frequency.y)),  # Color based on proportion, size based on total frequency
             alpha = 0.9) +  # Set transparency of the points
  
  # Comment out the following line if you don't want city labels
  # geom_text(data = merger, aes(x = lng, y = lat, label=city), hjust=0.1, vjust=0.1, size = 2) +
  
  # Define color gradient for the points from "seagreen3" to "mediumpurple3"
  scale_color_gradient(low = "seagreen3", 
                       high = "mediumpurple3") +
  
  # Remove the size legend (because size is based on the sum of frequencies, which may not need a legend)
  guides(size = "none") +
  
  # Add labels for the size legend, indicating it represents the 'Proportion'
  labs(size = "Proportion") +
  
  # Set the plot title to the newly defined 'newtitle' with italicized words
  ggtitle(newtitle) +
  
  # Customize the plot's theme to remove axis titles and text, clean up grid lines, and style the legend and title
  theme(axis.title.x = element_blank(),       # Remove x-axis title
        axis.title.y = element_blank(),       # Remove y-axis title
        axis.text.x = element_blank(),        # Remove x-axis text
        axis.text.y = element_blank(),        # Remove y-axis text
        panel.grid.major = element_blank(),   # Remove major grid lines
        legend.text = element_text(size = 6),       # Set the legend text size
        legend.title = element_text(size = 9),      # Set the legend title size
        plot.title = element_text(hjust = 0.4,      # Horizontally adjust the plot title
                                  size = 13,        # Set the plot title size
                                  vjust = -2))      # Vertically adjust the plot title
```
