---
title: "Moran's I"
author: "Dana Roemling"
date: '2023-05-25'
output: html_document
toc: true
toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r libs}
# Load necessary libraries
library(dplyr)   # For data manipulation
library(spdep)   # For spatial analysis
```

## Data 

```{r data}
# Read in the geographical coordinates of cities from a CSV file
longlat <- read.csv(file = './data_maps/gsa_geo_filtered.csv')

# Extract only longitude and latitude columns, ignoring city names
longlat_only <- longlat %>% dplyr::select(lon, lat)

# Read in linguistic data and token counts at different locations
corpus <- read.csv(file = './data_ling/full_matrix_for_filtering.csv')   # Full linguistic matrix
token_at_location <- read.csv(file = './data_ling/tokens_at_location.csv')  # Token counts by city
colnames(token_at_location) <- c("City", "Tokencount")   # Rename columns for clarity

# Extract data for a specific word from the corpus
one_word <- corpus %>% filter(word == "nix")   # Filter for the word "nix"
colnames(one_word) <- c("Token", "City", "Frequency")  # Rename columns for clarity

# Merge the filtered data with geographic data (adding longitude and latitude to each city)
merger_one <- merge(one_word, longlat, by.x = "City", by.y = "City")  # Merge on city names
colnames(merger_one) <- c("City", "Token", "Frequency", "lon", "lat")  # Rename columns after merge

# Merge the data again, this time adding token count data to calculate relative frequency
merger <- merge(merger_one, token_at_location, by.x = "City", by.y = "City")
merger$relfreq <- (merger$Frequency / merger$Tokencount)  # Calculate relative frequency per token count
merger$relfreq1000 <- (merger$relfreq * 1000)  # Scale relative frequency per 1000 tokens

# Rename columns for clarity
colnames(merger) <- c("City", "Token", "Frequency", "lon", "lat", "TokenCount", "RelativeFrequency", "RelativeFrequencyThousand")

# Exclude cities with token counts in the first quantile (low-frequency locations)
quant <- summary(merger$TokenCount)[2]  # Calculate the first quantile of token count
merger <- merger %>% filter(TokenCount > quant)  # Filter out cities with token counts <= first quantile

# Create a smaller version of the merged data with only relevant columns
merger_small <- merger %>% dplyr::select("City", "lon", "lat", "RelativeFrequency")

# Sort the smaller dataset by longitude in decreasing order
merger_small <- merger_small[order(merger_small$lon, decreasing=TRUE),]  # Order by longitude
```

## Resolve duplicate long/lat points

Given the above corpus search, resolve duplicate places by getting their mean (of relative frequency) using the aggregate function below.

```{r resolve duplicates}
# Perform spatial data analysis, which may fail if there are duplicate entries at the same location.
# To handle duplicates, sum their relative frequencies for those locations.
merger_single <- aggregate(merger_small$RelativeFrequency, 
                           by=list(lon=merger_small$lon, lat=merger_small$lat), 
                           FUN=mean)  # Aggregate by longitude and latitude, taking the mean of RelativeFrequency

# Reorder the merged data to maintain the same order based on longitude in decreasing order.
merger_single <- merger_single[order(merger_single$lon, decreasing=TRUE),] 

# Rename columns for clarity and readability, indicating that they represent coordinates and relative frequency.
colnames(merger_single) <- c("lon", "lat", "RelativeFrequency")  
```

## Moran's I

Check help(nb2listw) to see different styles for weighting. 

```{r moran}
# Extract only the longitude and latitude columns from the aggregated data for further spatial analysis.
longlat_variable <- merger_single %>% dplyr::select(lon, lat)

# Calculate the k-nearest neighbors (knn) using the geographical coordinates.
# If longlat = TRUE, Great Circle distances are used for calculating distances between points.
# More information on this can be found in the tutorial linked below.
# https://spatialanalysis.github.io/lab_tutorials/Distance_Based_Spatial_Weights.html
knn <- knn2nb(knearneigh(coordinates(longlat_variable), k = 10, longlat = TRUE))  # Get 10 nearest neighbors

# Include the self-neighbor in the knn list to ensure that each location is considered its own neighbor.
knn <- include.self(knn)

# Convert the knn object into a list of spatial weights.
# The weights are based on the neighbors found, and zero.policy = TRUE allows for handling cases with no neighbors.
# More about this process can be found in the following link.
# https://mgimond.github.io/simple_moransI_example/
weighted_knn <- nb2listw(knn, style = "W", zero.policy = TRUE)

# Calculate Moran's I statistic for spatial autocorrelation based on the relative frequencies.
# The function returns a vector, so we extract the first element which represents the Moran's I value.
# The formula used for the calculation is adapted from the provided example link.
moran <- moran(merger_single$RelativeFrequency, weighted_knn, length(knn), Szero(weighted_knn))[1]
```

It is important that all data inputted into the moran() is ordered the same way, because the order will determine alignment of variable data, location and neighbours. 
- merger & merger_small are ordered by city name. So I reorder by longitude, as this value is in most other data frames.
- merger_single is also reordered by longitude, just to be sure it is consistent. 
- longlat_variable is ordered like merger_single / uses same order
- knn seems plausible to have the same order as longlat_variable, however does not contain lon/lat, just index of numbers
- weighted_knn keeps the order from knn. 


```{r check knn order}
# Calculate the k-nearest neighbors (knn) for the geographical coordinates in longlat_variable.
# Here, k = 2 indicates that we are looking for the 2 nearest neighbors for each point.
# The longlat = TRUE argument ensures that Great Circle distances are used, which is appropriate for geographical coordinates.
# The knn2nb function converts the knn object into a neighbor list format suitable for spatial analysis.
knn_check <- knn2nb(knearneigh(coordinates(longlat_variable), k = 2, longlat = TRUE))  
```

As there is no long/lat in knn, so with 2 nearest, I can check for plausibility of nearest neighbours. 
The data is ordered by longitude, so it makes sense that now the neighbours can be close as well.
In this example case, Wien has the neighbours Wiener Neudorf and Leoben. Same for Wiender Neudorf itself. Stralsund neighbours are 14/24: Bentwisch and Greifswald. This seems plausible. 


## Just a reminder about Moran's I

https://en.wikipedia.org/wiki/Moran%27s_I
Values of I usually range from −1 to +1. Values significantly below -1/(N-1) indicate negative spatial autocorrelation and values significantly above -1/(N-1) indicate positive spatial autocorrelation.
Moran's I is inversely related to Geary's C, but it is not identical. Moran's I is a measure of global spatial autocorrelation, while Geary's C is more sensitive to local spatial autocorrelation.
