---
title: "Moran's I"
author: "Dana Roemling"
date: '2023-05-25'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r libs}
library(dplyr)
library(spdep)
```

## Data 

```{r data}
# get all cities in corpus and make geographical matrix out of it based on distance.
longlat <- read.csv(file = './data_maps/gsa_geo_filtered.csv')

# ignore city names
longlat_only <- longlat %>% dplyr::select(lon, lat)

# ling data
corpus <- read.csv(file = './data_ling/full_matrix_for_filtering.csv') 
token_at_location <- read.csv(file = './data_ling/tokens_at_location.csv')
colnames(token_at_location) <- c("City", "Tokencount") 

# extract feature
one_word <- corpus %>% filter(word == "nix")
colnames(one_word) <- c("Token", "City", "Frequency")
#merge with geo data
merger_one <- merge(one_word, longlat, by.x ="City", by.y = "City")
#merger_one_2 <- merge(one_word, longlat_small, by.x ="City", by.y = "City")
colnames(merger_one) <- c("City", "Token", "Frequency", "lon", "lat")
# add relative frequency
merger <- merge(merger_one, token_at_location, by.x ="City", by.y = "City")
merger$relfreq <- (merger$Frequency/merger$Tokencount)
merger$relfreq1000 <- (merger$relfreq*1000)
colnames(merger) <- c("City", "Token", "Frequency", "lon", "lat", "TokenCount", "RelativeFrequency", "RelativeFrequencyThousand")
# exclude first quantile / low frequency locations
quant <- summary(merger$TokenCount)[2]
merger <- merger %>% filter(TokenCount > quant)

# small version of merger
merger_small <- merger %>% dplyr::select("City", "lon", "lat", "RelativeFrequency")

# order by lon
merger_small <- merger_small[order(merger_small$lon,decreasing=TRUE),]
```

## Resolve duplicate long/lat points

Given the above corpus search, resolve duplicate places by getting their mean (of relative frequency) using the aggregate function below.

```{r resolve duplicates}
# Spatial data analysis may fail, so I need to sort my duplicates. Since they are at the same location,
# easiest would be to just sum them
merger_single <- aggregate(merger_small$RelativeFrequency, by=list(lon=merger_small$lon, lat=merger_small$lat), FUN=mean)
# reorder so we keep same order
merger_single <- merger_single[order(merger_single$lon,decreasing=TRUE),]
# readable names
colnames(merger_single) <- c("lon", "lat", "RelativeFrequency")
```

## Moran's I

Check help(nb2listw) to see different styles for weighting. 

```{r moran}
# get only long lat
longlat_variable <- merger_single %>% dplyr::select(lon, lat)
# get knn
# If longlat = TRUE, Great Circle distances are used
# https://spatialanalysis.github.io/lab_tutorials/Distance_Based_Spatial_Weights.html
knn <- knn2nb(knearneigh(coordinates(longlat_variable), k = 10, longlat = TRUE))
# include self
knn <- include.self(knn)
# assign weights
# https://mgimond.github.io/simple_moransI_example/
weighted_knn <- nb2listw(knn, style="W", zero.policy=TRUE)
# calculate moran
moran <- moran(merger_single$RelativeFrequency, weighted_knn, length(knn), Szero(weighted_knn))[1]
# formula adapted from https://mgimond.github.io/simple_moransI_example/
```

It is important that all data inputted into the moran() is ordered the same way, because the order will determine alignment of variable data, location and neighbours. 
- merger & merger_small are ordered by city name. So I reorder by longitude, as this value is in most other data frames.
- merger_single is also reordered by longitude, just to be sure it is consistent. 
- longlat_variable is ordered like merger_single / uses same order
- knn seems plausible to have the same order as longlat_variable, however does not contain lon/lat, just index of numbers
- weighted_knn keeps the order from knn. 


```{r check knn order}
knn_check <- knn2nb(knearneigh(coordinates(longlat_variable), k = 2, longlat = TRUE))
```

As there is no long/lat in knn, so with 2 nearest, I can check for plausibility of nearest neighbours. 
The data is ordered by longitude, so it makes sense that now the neighbours can be close as well.
In this example case, Wien has the neighbours Wiener Neudorf and Leoben. Same for Wiender Neudorf itself. Stralsund neighbours are 14/24: Bentwisch and Greifswald. This seems plausible. 


## Just a reminder about Moran's I

https://en.wikipedia.org/wiki/Moran%27s_I
Values of I usually range from −1 to +1. Values significantly below -1/(N-1) indicate negative spatial autocorrelation and values significantly above -1/(N-1) indicate positive spatial autocorrelation.
Moran's I is inversely related to Geary's C, but it is not identical. Moran's I is a measure of global spatial autocorrelation, while Geary's C is more sensitive to local spatial autocorrelation.




