---
title: "Kriging"
author: "Dana Roemling"
date: '2024-03-07'
output: html_document
toc: true
toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Kriging

In this markdown I will save all the code I need for the kriging and the interpolation of maps.


### Preparation

I'll start with the setup.

```{r libs_data}
library(sf)
library(sp)
library(tidyverse)
library(gstat)
library(stringr)
library(dplyr) 
library(tidyr)
library(scales)
library(classInt) 
library(rnaturalearth)
library(rnaturalearthhires) # this is tricky to install
# you can just use rnaturalearth and then set scale = "small" in the ne_countries() function

# data on local
corpus <- read.csv(file = './data_ling/full_matrix_for_filtering.csv') 
token_at_location <- read.csv(file = './data_ling/tokens_at_location.csv')
colnames(token_at_location) <- c("City", "Tokencount") 
longlat <- read.csv(file = './data_maps/gsa_geo_filtered.csv')
longlat_small <- read.csv(file = './data_maps/gsa_geo_filtered_nosmall22.csv')
```

I'll also create an object with some cities, which I can later use to orientate people on the map.
This is not needed for the kriging itself, just to make the map prettier.

```{r cities}
cities <- data.frame(
  City = c("Cologne", "Munich", "Vienna", "Zurich", "Berlin", "Hamburg"),
  Long = c(6.9578, 11.5755, 16.3731, 8.5417, 13.3833, 10),
  Lat = c(50.9422, 48.1372, 48.2083, 47.3769, 52.5167, 53.55))
crs2 <- CRS("+init=epsg:4326")
cities_sf <- st_as_sf(cities, coords = c("Long", "Lat"), crs = crs2)
```

This creates the grids we need for kriging. ne_countries should have the world, so it should be easily adjustable to map other areas. 

```{r grid}
gsa_outline <- ne_countries(country = c("Austria", "Germany", "Switzerland"), returnclass="sf", scale = "large")
gsa_plot <- gsa_outline %>% dplyr::select(geometry) # just the outlines of the three
gsa_spatial <- as_Spatial(gsa_plot) # new object as spatial instead of sf

# large grid
sp_grid <- as.data.frame(spsample(gsa_spatial,
                                  n = 100000,
                                  type = "regular", # systematically aligned sampling
                                  offset = c(0.5,0.5))) # makes sure grid is the same every time
# This grid needs A CRS and to be turned into a spatial object, which this line does:
sp_grid_sf <- st_as_sf(sp_grid, coords=c("x1","x2"), crs = st_crs(4326))



# small grid
sp_grid_small <- as.data.frame(spsample(gsa_spatial,
                                  n = 500,
                                  type = "regular", # systematically aligned sampling
                                  offset = c(0.5,0.5))) # makes sure grid is the same every time
# This grid needs A CRS and to be turned into a spatial object, which this line does:
sp_grid_sf_small <- st_as_sf(sp_grid_small, coords=c("x1","x2"), crs = st_crs(4326))
```

### Data

Finding what I actually like to krige.

```{r counting / regex}
#get word of interest
one_word <- corpus %>% filter(word == "guck")
colnames(one_word) <- c("Token", "City", "Frequency")

# join to all longlat and introduce 0 for NA
new_join <- merge(longlat_small, one_word, by.x ="City", by.y = "City", all = TRUE)
new_join <- new_join %>% dplyr::select("City", "lon", "lat", "Frequency")
new_join[is.na(new_join)] <- 0

# OR

#merge with geo data
merger_one <- merge(one_word, longlat, by.x ="City", by.y = "City")
colnames(merger_one) <- c("City", "Token", "Frequency", "lon", "lat")

# add relative frequency
merger <- merge(merger_one, token_at_location, by.x ="City", by.y = "City")
merger$relfreq <- (merger$Frequency/merger$Tokencount)
merger$relfreq1000 <- (merger$relfreq*1000)
colnames(merger) <- c("City", "Token", "Frequency", "lon", "lat", "TokenCount", "RelativeFrequency", "RelativeFrequencyThousand")
summary(merger$TokenCount)
merger <- merger %>% filter(TokenCount > 4000)
```


### Variograms

Now that I have the grid ready, I can get to the next step. For the kriging to work, I need to create Variograms first. A variogram is 'a function that, given a distance between two observations, returns the variance of the difference between those observations'. Simpler put, it 'describes the amount of spatial variability in the observed values'.

For that to happen, I need my data. From this data, I create a Spatial Points Data Frame, which is what gstats needs in order to produce a variogram. This df contains projected coordinates. I could also create a SPDF by using coordinates(), but this wouldn't project the data.

The variogram function takes an object, the data and several other arguments. The object needs to be a formula. In the case of a missing regressor, we use a constant mean model denoted by ~1 after the variable of interest. The data needs to contain the spatial information of the variable. 

Cutoff: as a default, the length of the diagonal of the box spanning the data is divided by three.
Width: Bin/Interval in the distance used by the model. 
cloud = TRUE creates an empirical variogram / adds the cloud. Otherwise a theoretical variogram. 

For projected data, Euclidian distances are computed, for unprojected great circle distances (km).

This code outputs a variogram with fit line, which I then can use for the kriging. 

```{r single_value_variogram}
# create the data frame needed for variogram
trial <- merger %>% dplyr::select("lon", "lat", "RelativeFrequencyThousand") #%>% na.omit()
data <- trial[3]
coords <- trial[1:2]
crs <- CRS("+init=epsg:4326") # Different way of setting same CRS
# it will give a warning, but work nontheless
#crs <- CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")
spdf <- SpatialPointsDataFrame(coords      = coords,
                               data        = data, 
                               proj4string = crs)

#need to remove duplicates
spdf = spdf[which(!duplicated(spdf@coords)), ]

# variogram
vg <- variogram(RelativeFrequencyThousand~1, data = spdf, width = 1, cutoff = 300)
vg_fit <- fit.variogram(vg, vgm("Exp")) 
plot(vg, vg_fit)
```

### Actual Kriging

The actual kriging needs three inputs: 
- The original data
- The grid
- The fit from the variogram, represented by nugget, sill and range values

Similar to the variogram, I pass the variable I am interested in, Frequency, as part of a function with a constant mean model. I then pass the original data and the grid I produced. The last argument is the model I choose for the kriging. 

```{r kriging}
krig1 <- krige(RelativeFrequencyThousand~1, 
               locations = spdf, 
               newdata = sp_grid_sf, 
               model = vg_fit)
```

I can also run the kriging twice, to smooth out the data more. For this, I first run the kriging on a small grid, like 50-500 cells. Then I use that and input it into a second kriging.

```{r smoothing_kriging}
krig1 <- krige(RelativeFrequencyThousand~1, 
               locations = spdf, 
               newdata = sp_grid_sf_small, 
               model = vg_fit)

krig2 <- krige(var1.pred~1, 
               locations = krig1, 
               newdata = sp_grid_sf, 
               model = vg_fit)
```

### Plotting kriging / point plots

As the last step I can then make a pretty map based on the kriging. First I set up the legend text.

```{r legend_text}
legend_max <- round(max(krig3$var1.pred), digits = 2)
legend_min <- round(min(krig3$var1.pred), digits = 2)
```

This creates the actual plot.

```{r lightmode_plot_purple_nbands}
ggplot() +
  geom_sf(data = krig3, aes(fill=var1.pred), shape = 21, size = 0.5, stroke = 0, lwd = 0) +
  geom_sf(data = gsa_plot, aes(geometry = geometry), color="black", fill=NA, size = 0.5) +
  geom_sf_text(data = cities_sf, aes(label = City), size = 2.5, nudge_x = 0, nudge_y = -0.15, family = "Optima") +
  geom_sf(data = cities_sf, aes(geometry = geometry), shape = 4) +
  theme_minimal() +
  scale_fill_steps(low="#ffffff", high="#310d59", n.breaks = 30, labels = NULL) +
  theme(axis.title.x = element_blank(), 
        axis.title.y = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        panel.grid.major = element_blank()) +
  theme(legend.position = c(0.90, 0.65), 
        legend.title = element_text(size= 6, family = "Optima"), 
        legend.text = element_text(size = 6, family = "Optima", hjust = 1),
        legend.title.align = 0,
        legend.key.size = unit(0.3, "cm"),
        legend.key.width = unit(0.4,"cm")) +
  annotate(geom="text", x = 16.6, y = 52.4, label=legend_max, size = 1.5, family = "Optima") +
  annotate(geom="text", x = 16.6, y = 51.3, label=legend_min, size = 1.5, family = "Optima") +
  labs(fill = "Mehr 'nichts'") 
```

### Exporting

To save files, I can use jpeg() or tiff().

```{r saving_maps}
tiff("NAME.tiff", units="in", width=3.78, height=4, res=300)
# OR
# jpeg("NAME.jpg", units="in", width=3.5, height=2, res=300)

# DELETE THIS AND HAVE PLOT CODE HERE

dev.off()
```